---
title: "WHI Abuse and Sexual Function"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
---

```{r setup, include=FALSE}
# Load packages
library(flexdashboard)
library(tidyverse)
library(ggrepel)
```

```{r load_data}
baseline_only <- read_rds(path = "../data/baseline_only.rds")
no_abuse <- read_rds(path = "../data/no_abuse.rds")
any_abuse <- read_rds(path = "../data/any_abuse.rds")
verbal_only <- read_rds(path = "../data/verbal_only.rds")
physical_only <- read_rds(path = "../data/physical_only.rds")
verbal_and_physical <- read_rds(path = "../data/verbal_and_physical.rds")
table_01_n <- nrow(baseline_only) %>% format(big.mark = ",")
```

Sidebar {.sidebar data-width=300}
===============================================================================
All analyses include information from women who participated in the [WHI](https://www.whi.org/SitePages/WHI%20Home.aspx), gave a "yes" or "no" response to the two questions related to abuse, and reported being sexually active in the previous year.

Please use the navigation bar (blue bar at the top of the screen) to view:  

1. **Table 1.** Sociodemographic and health history characteristics of women participating in the WHI by abuse experience at baseline  

2. **Figure 1.** Percent of Women Who Reported Abuse at Baseline by Age Group    

3. **Figure 2.** Percent of Women Who Reported Sexual Satisfaction at Baseline by Age Group   

4. **Figure 3.** Percent of Women Who Reported Sexual Frequency Satisfaction at Baseline by Age Group

5. **Table 2.** Unadjusted and adjusted odds of global sexual dissatisfaction and sexual frequency dissatisfaction at baseline 

6. **Table 3.** Number and percent of women who report sexual satisfaction and sexual frequency satisfaction at baseline by age group and abuse status

Overview
===============================================================================

Row 1
-------------------------------------------------------------------------------

### Reported No Abuse at Baseline

```{r}
n <- nrow(no_abuse) %>% format(., big.mark = ",")
valueBox(n, color = "#63A2E7")
```

### Reported Any Abuse at Baseline

```{r}
n <- nrow(any_abuse) %>% format(., big.mark = ",")
valueBox(n, color = "#CC0000")
# Add color
```

### Reported Verbal Abuse Only at Baseline

```{r}
n <- nrow(verbal_only) %>% format(., big.mark = ",")
valueBox(n, color = "#CC0000")
```

### Reported Physical Abuse Only at Baseline

```{r}
n <- nrow(physical_only) %>% format(., big.mark = ",")
valueBox(n, color = "#CC0000")
```

### Reported Physical and Verbal Abuse at Baseline

```{r}
n <- nrow(verbal_and_physical) %>% format(., big.mark = ",")
valueBox(n, color = "#CC0000")
```

Row 2
-------------------------------------------------------------------------------

### Sexual Satisfaction by Reported Abuse at Baseline

```{r}
baseline_only %>% 
  group_by(abuse_d_f, satisfied_f) %>% 
  count() %>% 
  mutate(
    cumsum  = cumsum(n),
    N = max(cumsum),
    prop = n / N,
    percent = round((prop * 100), 1),
    se_of_prop = sqrt(prop * (1-prop)/(N-1)),
    tstat = Vectorize(qt, "df")(0.975, N-1),
    lower = (prop - tstat * se_of_prop) * 100,
    upper = (prop + tstat * se_of_prop) * 100
  ) %>% 
  ggplot(data = ., aes(x = satisfied_f, y = percent)) +
    geom_point(aes(color = abuse_d_f), stat = "identity", size = 3, alpha = 0.7) +
    geom_label_repel(aes(label = percent, color = abuse_d_f), nudge_x = 0.25, show.legend = FALSE) +
    scale_x_discrete("Satisfied with Current Sexual Activities") +
    scale_y_continuous("Percent", limits = c(0, 100)) +
    scale_color_manual(values = c("#63A2E7", "#CC0000"), name = "Abuse at Baseline") +
    ggtitle("Sexual Satisfaction by Reported Abuse at Baseline") +
    theme_bw()
```

### Sexual Satisfaction by Reported Abuse at Baseline

```{r}
baseline_only %>% 
  group_by(abuse4cat_f, satisfied_f) %>% 
  count() %>% 
  mutate(
    cumsum  = cumsum(n),
    N = max(cumsum),
    prop = n / N,
    percent = round((prop * 100), 1),
    se_of_prop = sqrt(prop * (1-prop)/(N-1)),
    tstat = Vectorize(qt, "df")(0.975, N-1),
    lower = (prop - tstat * se_of_prop) * 100,
    upper = (prop + tstat * se_of_prop) * 100
  ) %>% 
  ggplot(data = ., aes(x = satisfied_f, y = percent)) +
    geom_point(aes(color = abuse4cat_f), stat = "identity", size = 3, alpha = 0.7) +
    geom_label_repel(aes(label = percent, color = abuse4cat_f), nudge_x = 0.25, show.legend = FALSE) +
    scale_x_discrete("Satisfied with Current Sexual Activities") +
    scale_y_continuous("Percent", limits = c(0, 100)) +
    scale_color_manual(values = c("#63A2E7", "#EDC33B", "#F46D33", "#CC0000"), name = "Abuse at Baseline") +
    ggtitle("Sexual Satisfaction by Reported Abuse at Baseline") +
    theme_bw()
```

Row 3
-------------------------------------------------------------------------------

### Sexual Frequency Satisfaction by Reported Abuse at Baseline

```{r}
baseline_only %>% 
  group_by(abuse_d_f, freq_satisfied_f) %>% 
  count() %>% 
  mutate(
    cumsum  = cumsum(n),
    N = max(cumsum),
    prop = n / N,
    percent = round((prop * 100), 1),
    se_of_prop = sqrt(prop * (1-prop)/(N-1)),
    tstat = Vectorize(qt, "df")(0.975, N-1),
    lower = (prop - tstat * se_of_prop) * 100,
    upper = (prop + tstat * se_of_prop) * 100
  ) %>% 
  ggplot(data = ., aes(x = freq_satisfied_f, y = percent)) +
    geom_point(aes(color = abuse_d_f), stat = "identity", size = 3, alpha = 0.7) +
    geom_label_repel(aes(label = percent, color = abuse_d_f), nudge_x = 0.25, show.legend = FALSE) +
    scale_x_discrete("Satisfied with Current Sexual Activities") +
    scale_y_continuous("Percent", limits = c(0, 100)) +
    scale_color_manual(values = c("#63A2E7", "#CC0000"), name = "Abuse at Baseline") +
    ggtitle("Sexual Frequency Satisfaction by Reported Abuse at Baseline") +
    theme_bw()
```

### Sexual Frequency Satisfaction by Reported Abuse at Baseline

```{r}
baseline_only %>% 
  group_by(abuse4cat_f, freq_satisfied_f) %>% 
  count() %>% 
  mutate(
    cumsum  = cumsum(n),
    N = max(cumsum),
    prop = n / N,
    percent = round((prop * 100), 1),
    se_of_prop = sqrt(prop * (1-prop)/(N-1)),
    tstat = Vectorize(qt, "df")(0.975, N-1),
    lower = (prop - tstat * se_of_prop) * 100,
    upper = (prop + tstat * se_of_prop) * 100
  ) %>% 
  ggplot(data = ., aes(x = freq_satisfied_f, y = percent)) +
    geom_point(aes(color = abuse4cat_f), stat = "identity", size = 3, alpha = 0.7) +
    geom_label_repel(aes(label = percent, color = abuse4cat_f), nudge_x = 0.25, show.legend = FALSE) +
    scale_x_discrete("Satisfied with Current Sexual Activities") +
    scale_y_continuous("Percent", limits = c(0, 100)) +
    scale_color_manual(values = c("#63A2E7", "#EDC33B", "#F46D33", "#CC0000"), name = "Abuse at Baseline") +
    ggtitle("Sexual Frequency Satisfaction by Reported Abuse at Baseline") +
    theme_bw()
```

Table 1
===============================================================================

Row 1 {data-height=10}
-------------------------------------------------------------------------------

### Download in Word Format

[Click here to download Table 1 in Word Format](https://github.com/brad-cannell/whi_sexual_function/blob/master/R%20notebooks/table_1_word.docx)

Row 2
-------------------------------------------------------------------------------

### Table 1 Sociodemographic and health history characteristics of women participating in the WHI by abuse experience at baseline (N = `r table_01_n`)<sup>1</sup>

```{r table_01_kable}
table_01_kable <- read_rds("table_01_kable.rds") # Created in table_1_word.Rmd
table_01_kable
```

**Footnotes:**    
1. Baseline data for women who had any sexual activity with a partner in the previous year and had a known abuse status. Individual group n's are not mutually exclusive.    
2. All percentages above are column percentages. For example, 42% of women who reported no abuse at baseline were in the <50 to 59 age group at baseline.   

Figure 1
===============================================================================

### Percent, and 95% confidence interval, of Women Who Reported Abuse at Baseline by Age Group

```{r}
baseline_only %>% 
  group_by(age_group_f, abuse_d_f) %>% 
  count() %>% 
  mutate(
    cumsum  = cumsum(n),
    N = max(cumsum),
    prop = n / N,
    percent = round((prop * 100), 1),
    se_of_prop = sqrt(prop * (1-prop)/(N-1)),
    tstat = Vectorize(qt, "df")(0.975, N-1),
    lower = (prop - tstat * se_of_prop) * 100,
    upper = (prop + tstat * se_of_prop) * 100
  ) %>%
  filter(abuse_d_f == "Yes") %>% 
  ggplot(data = ., aes(x = age_group_f, y = percent)) +
    geom_point(color = "blue", stat = "identity", size = 3) +
    geom_errorbar(aes(x = age_group_f, ymin = lower, ymax = upper), width = 0.15) +
    geom_label_repel(aes(label = percent), nudge_x = 0.25, show.legend = FALSE) +
    scale_x_discrete("Age Group at Baseline") +
    scale_y_continuous("Percent", limits = c(0, 25)) +
    theme_bw()
```

Figure 2
===============================================================================

### Percent, and 95% confidence interval, of Women Who Reported Sexual Satisfaction at Baseline by Age Group

```{r}
baseline_only %>% 
  group_by(age_group_f, satisfied_f) %>% 
  count() %>% 
  mutate(
    cumsum  = cumsum(n),
    N = max(cumsum),
    prop = n / N,
    percent = round((prop * 100), 1),
    se_of_prop = sqrt(prop * (1-prop)/(N-1)),
    tstat = Vectorize(qt, "df")(0.975, N-1),
    lower = (prop - tstat * se_of_prop) * 100,
    upper = (prop + tstat * se_of_prop) * 100
  ) %>%
  filter(satisfied_f == "Yes") %>% 
  ggplot(data = ., aes(x = age_group_f, y = percent)) +
    geom_point(color = "blue", stat = "identity", size = 3) +
    geom_errorbar(aes(x = age_group_f, ymin = lower, ymax = upper), width = 0.15) +
    geom_label_repel(aes(label = percent), nudge_x = 0.25, show.legend = FALSE) +
    scale_x_discrete("Age Group at Baseline") +
    scale_y_continuous("Percent", limits = c(0, 100)) +
    theme_bw()
```

Figure 3
===============================================================================

### Percent, and 95% confidence interval, of Women Who Reported Sexual Frequency Satisfaction at Baseline by Age Group

```{r}
baseline_only %>% 
  filter(!is.na(satfrqsx_f)) %>% 
  group_by(age_group_f, satfrqsx_f) %>% 
  count() %>% 
  mutate(
    cumsum  = cumsum(n),
    N = max(cumsum),
    prop = n / N,
    percent = round((prop * 100), 1),
    se_of_prop = sqrt(prop * (1-prop)/(N-1)),
    tstat = Vectorize(qt, "df")(0.975, N-1),
    lower = (prop - tstat * se_of_prop) * 100,
    upper = (prop + tstat * se_of_prop) * 100
  ) %>% 
  ggplot(data = ., aes(x = age_group_f, y = percent)) +
    geom_point(aes(color = satfrqsx_f), stat = "identity", size = 3) +
    geom_errorbar(aes(x = age_group_f, ymin = lower, ymax = upper, color = satfrqsx_f), width = 0.15) +
    geom_label_repel(aes(label = percent, color = satfrqsx_f), nudge_x = 0.25, show.legend = FALSE) +
    scale_x_discrete("Age Group at Baseline") +
    scale_y_continuous("Percent", limits = c(0, 100)) +
    scale_color_manual(values = c("#EDC33B", "#F46D33", "#CC0000"), name = "Sexual Freq Satisfaction") +
    theme_bw()
```


Table 2
===============================================================================

Row 1 {data-height=10}
-------------------------------------------------------------------------------

### Download in Word Format

[Click here to download Table 2 in Word Format](https://github.com/brad-cannell/whi_sexual_function/blob/master/R%20notebooks/table_2_word.docx)

Row 2
-------------------------------------------------------------------------------

### **Table 2.** Unadjusted and adjusted<sup>1</sup> odds of global sexual dissatisfaction<sup>2</sup> and sexual frequency dissatisfaction<sup>3</sup> at baseline<sup>4</sup> (N = 83,218). 

```{r table_02_kable}
table_02_kable <- read_rds("table_02_kable.rds") # Created in table_2_word.Rmd
table_02_kable
```

**Footnotes:**  
1. Adjusted for age, race/ethnicity education income, marital status, sexual orientation, study component, parity, alcohol consumption, caffeine consumption, smoking status, hormone use, use of SSRIs, quality of life, depression, BMI, self-rated health, hysterectomy, night sweats, hot flashes, vaginal dryness, urinary incontinence, chronic disease.
2. How satisfied are you with your current sexual activities, either with a partner or alone?   
3. Are you satisfied with the frequency of your sexual activity, or would you like to have sex more or
less often?   
4. Baseline data for women who had any sexual activity with a partner in the previous year, had a known abuse status, and responded to the sexual satisfaction questions.    

Table 3
===============================================================================

Row 1 {data-height=10}
-------------------------------------------------------------------------------

### Download in Word Format

[Click here to download Table 3 in Word Format](https://github.com/brad-cannell/whi_sexual_function/blob/master/R%20notebooks/table_3_word.docx)

Row 2 
-------------------------------------------------------------------------------
### **Table 3.** Number and percent of women who report sexual satisfaction<sup>1</sup> and sexual frequency satisfaction<sup>2</sup> at baseline by age group and abuse status, results from the Women’s Health Initiative (WHI) (N = 78,416)<sup>3</sup>. 

```{r table_03_kable}
table_03_kable <- read_rds("table_03_kable.rds") # Created in table_3_word.Rmd
table_03_kable
```

**Footnotes:**   
1. How satisfied are you with your current sexual activities, either with a partner or alone?   
2. Are you satisfied with the frequency of your sexual activity, or would you like to have sex more or
less often?   
3. Baseline data for women who had any sexual activity with a partner in the previous year, had a known abuse status, and responded to the sexual satisfaction questions.   
4. _Desire greater sexual frequency_ is a subset of women who were _unsatisfied with sexual frequency_. The percent given is the row percent. For example, 27.4% of women who reported no abuse at baseline reported a desire for sexual frequency more often.   
