---
title: "Exploring logistic models before creating Table 2"
date: "Created: 2017-05-29 <br> Updated: `r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages,include=FALSE}
# Load packages
library(tidyverse)
library(feather)
library(bfuncs)
library(mice)
library(broom)
library(modelr)
```

```{r load_data, include=FALSE}
# Load data
mids_83329 <- read_rds(path = "../data/mids_83329.rds")
analysis_11 <- read_feather("../data/analysis_11.feather")
```

```{r}
about_data(analysis_11) # 916,619 observations and 59 variables
```


From Erika 2017-07-17: 

## Step 2: Compute odds of sexual satisfaction (Table 2). 

Conduct adjusted models to examine cross-sectional odds of sexual satisfaction (Global/frequency), with sequential modeling as described below. Use multiple imputation to account for the missingness of chronic disease information. Sequential modeling as follows:

**Model 1 (Sociodemographics):**   
* Age   
* Race & Ethnicity   
* Education   
* Income   
* Marital Status   
* WHI study component   
	
**Model 2 (Health/health risk factors + Model 1):**   
* All Model 1 variables   
* Pack years of smoking   
* Body mass index (BMI)   
* Hysterectomy ever   
* Urinary incontinence ever   
* Hypertension ever (variable label = hypt; self-report Form 30)   
* Cardiovascular disease ever (variable label = CVD; self-report Form 30)   
* Arthritis ever (Label = arthrit; self-report Form 30)   
* Cancer ever (Label = canc_f30; Form 30)   
* Diabetes ever (Variable label = diab & diabtrt; self-report Form 2)   
* Hip fracture age 55 or older (Variable label = hip55; self-report Form 30)   
* Self-rated health   
* Depressive symptoms   
* Current use of SSRI   

**Model 3 (Menopausal symptoms/treatment variables that impact menopausal + Model 1 +Model 2):**   
* All Model 1 variables   
* All Model 2 variables   
* Night sweats   
* Hot flashes   
* Vaginal dryness   
* Current use of female hormone therapy   

### Let's explore the variable distributions by imputation status

What does the distribution of satisfied by abuse look like?

```{r}
analysis_11 %>% 
  filter(!.imp == 0) %>% 
  group_by(abuse, satisfied_f) %>% 
  freq_table() %>% 
  select(1:3, percent_row, lcl_row_log, ucl_row_log) %>% 
  filter(satisfied_f == "Yes")
```

```{r}
analysis_11 %>% 
  filter(!.imp == 0) %>%
  group_by(abuse, freq_satisfied_f) %>% 
  freq_table() %>% 
  select(1:3, percent_row, lcl_row_log, ucl_row_log)
```

## Using mice::with with return a Multiply imputed repeated analyses (MIRA) object

Create helper function for displaying model output

```{r}
tidy_output <- function(x, ...) {
  x %>% 
    summary() %>% 
    tidy() %>% 
    mutate(
    est = exp(est) %>% round(2) %>% format(nsmall = 2L),
    lo.95 = exp(lo.95) %>% round(2) %>% format(nsmall = 2L),
    hi.95 = exp(hi.95) %>% round(2) %>% format(nsmall = 2L),
    or_95 = paste0(est, " (", lo.95, " - ", hi.95, ")")
  ) %>% 
  select(.rownames, est, lo.95, hi.95, or_95)
}
```










## Unadjusted model dissatisfaction

-------------------------------------------------------------------------------

```{r}
unadj_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse,
    family = binomial(link = "logit")
  )
)
```

```{r}
unadj_dissat_pool <- pool(unadj_dissat)
unadj_dissat_pool %>% tidy_output()
```

## Check for age group interaction

```{r}
analysis_10 <- read_feather("../data/analysis_10.feather")
baseline_only <- analysis_10 %>% filter(baseline == TRUE)
```

```{r}
interaction <- glm(
  satisfied_f ~ abuse_d_f + age + abuse_d_f:age,
  family = binomial(link = "logit"),
  data = analysis_10
)

interaction %>% summary()
confint(interaction) %>% exp
exp(interaction$coefficients)
```


```{r}
interaction <- pool(interaction)
interaction %>% tidy_output()
```

## Adjusted model 1 dissatisfaction

* Age   
* Race & Ethnicity   
* Education   
* Income   
* Marital Status   
* WHI study component   

```{r}
m1_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m1_dissat_pool <- pool(m1_dissat)
m1_dissat_pool %>% tidy_output()
```

## Adjusted model 2 dissatisfaction

* All Model 1 variables   
* Pack years of smoking   
* Body mass index (BMI)   
* Hysterectomy ever   
* Urinary incontinence ever   
* Hypertension ever (variable label = hypt; self-report Form 30)   
* Cardiovascular disease ever (variable label = CVD; self-report Form 30)   
* Arthritis ever (Label = arthrit; self-report Form 30)   
* Cancer ever (Label = canc_f30; Form 30)   
* Diabetes ever (Variable label = diab & diabtrt; self-report Form 2)   
* Hip fracture age 55 or older (Variable label = hip55; self-report Form 30)   
* Self-rated health   
* Depressive symptoms   
* Current use of SSRI  

```{r}
m2_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m2_dissat_pool <- pool(m2_dissat)
m2_dissat_pool %>% tidy_output()
```

## Adjusted model 3 dissatisfaction

* All Model 1 variables   
* All Model 2 variables   
* Night sweats   
* Hot flashes   
* Vaginal dryness   
* Current use of female hormone therapy 

```{r}
m3_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f + night_sweats_f + hot_flashes_f + vag_dry_f +
      hormnw_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m3_dissat_pool <- pool(m3_dissat)
m3_dissat_pool %>% tidy_output()
```










## Unadjusted model frequency dissatisfaction

-------------------------------------------------------------------------------

```{r}
unadj_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse,
    family = binomial(link = "logit")
  )
)
```

```{r}
unadj_freq_dissat_pool <- pool(unadj_freq_dissat)
unadj_freq_dissat_pool %>% tidy_output()
```

## Adjusted model 1 frequency dissatisfaction

* Age   
* Race & Ethnicity   
* Education   
* Income   
* Marital Status   
* WHI study component   

```{r}
m1_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m1_freq_dissat_pool <- pool(m1_freq_dissat)
m1_freq_dissat_pool %>% tidy_output()
```

## Adjusted model 2 frequency dissatisfaction

* All Model 1 variables   
* Pack years of smoking   
* Body mass index (BMI)   
* Hysterectomy ever   
* Urinary incontinence ever   
* Hypertension ever (variable label = hypt; self-report Form 30)   
* Cardiovascular disease ever (variable label = CVD; self-report Form 30)   
* Arthritis ever (Label = arthrit; self-report Form 30)   
* Cancer ever (Label = canc_f30; Form 30)   
* Diabetes ever (Variable label = diab & diabtrt; self-report Form 2)   
* Hip fracture age 55 or older (Variable label = hip55; self-report Form 30)   
* Self-rated health   
* Depressive symptoms   
* Current use of SSRI  

```{r}
m2_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m2_freq_dissat_pool <- pool(m2_freq_dissat)
m2_freq_dissat_pool %>% tidy_output()
```

## Adjusted model 3 frequency dissatisfaction

* All Model 1 variables   
* All Model 2 variables   
* Night sweats   
* Hot flashes   
* Vaginal dryness   
* Current use of female hormone therapy 

```{r}
m3_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f + night_sweats_f + hot_flashes_f + vag_dry_f +
      hormnw_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m3_freq_dissat_pool <- pool(m3_freq_dissat)
m3_freq_dissat_pool %>% tidy_output()
```









-------------------------------------------------------------------------------

## Step 3: Investigation of interaction – does abuse type matter?   
For the full sample, investigate whether there is an interaction of abuse type on sexual satisfaction. This will inform how we do subsidiary analyses/format what was Table 3.

```{r}
unadj_dissat_interaction <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse4cat_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
unadj_dissat_interaction <- pool(unadj_dissat_interaction)
unadj_dissat_interaction %>% tidy_output()
```

```{r}
levels(analysis_11$abuse4cat_f)
```


```{r}
analysis_11 %>% 
  filter(!.imp == 0) %>% 
  group_by(abuse4cat_f, dissatisfied) %>% 
  freq_table() %>% 
  select(1:3, percent_row, lcl_row_log, ucl_row_log) %>% 
  filter(dissatisfied == 1)
```

Are verbal abuse only, physical abuse only, and both different from each other?

## Adjusted model 1 dissatisfaction

* Age   
* Race & Ethnicity   
* Education   
* Income   
* Marital Status   
* WHI study component   

```{r}
m1_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse4cat_f + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m1_dissat_pool <- pool(m1_dissat)
m1_dissat_pool %>% tidy_output()
```

## Adjusted model 2 dissatisfaction

* All Model 1 variables   
* Pack years of smoking   
* Body mass index (BMI)   
* Hysterectomy ever   
* Urinary incontinence ever   
* Hypertension ever (variable label = hypt; self-report Form 30)   
* Cardiovascular disease ever (variable label = CVD; self-report Form 30)   
* Arthritis ever (Label = arthrit; self-report Form 30)   
* Cancer ever (Label = canc_f30; Form 30)   
* Diabetes ever (Variable label = diab & diabtrt; self-report Form 2)   
* Hip fracture age 55 or older (Variable label = hip55; self-report Form 30)   
* Self-rated health   
* Depressive symptoms   
* Current use of SSRI  

```{r}
m2_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse4cat_f + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m2_dissat_pool <- pool(m2_dissat)
m2_dissat_pool %>% tidy_output()
```

## Adjusted model 3 dissatisfaction

* All Model 1 variables   
* All Model 2 variables   
* Night sweats   
* Hot flashes   
* Vaginal dryness   
* Current use of female hormone therapy 

```{r}
m3_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse4cat_f + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f + night_sweats_f + hot_flashes_f + vag_dry_f +
      hormnw_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m3_dissat_pool <- pool(m3_dissat)
m3_dissat_pool %>% tidy_output()
```










## Unadjusted model frequency dissatisfaction

-------------------------------------------------------------------------------

```{r}
unadj_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse4cat_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
unadj_freq_dissat_pool <- pool(unadj_freq_dissat)
unadj_freq_dissat_pool %>% tidy_output()
```

## Adjusted model 1 frequency dissatisfaction

* Age   
* Race & Ethnicity   
* Education   
* Income   
* Marital Status   
* WHI study component   

```{r}
m1_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse4cat_f + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m1_freq_dissat_pool <- pool(m1_freq_dissat)
m1_freq_dissat_pool %>% tidy_output()
```

## Adjusted model 2 frequency dissatisfaction

* All Model 1 variables   
* Pack years of smoking   
* Body mass index (BMI)   
* Hysterectomy ever   
* Urinary incontinence ever   
* Hypertension ever (variable label = hypt; self-report Form 30)   
* Cardiovascular disease ever (variable label = CVD; self-report Form 30)   
* Arthritis ever (Label = arthrit; self-report Form 30)   
* Cancer ever (Label = canc_f30; Form 30)   
* Diabetes ever (Variable label = diab & diabtrt; self-report Form 2)   
* Hip fracture age 55 or older (Variable label = hip55; self-report Form 30)   
* Self-rated health   
* Depressive symptoms   
* Current use of SSRI  

```{r}
m2_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse4cat_f + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m2_freq_dissat_pool <- pool(m2_freq_dissat)
m2_freq_dissat_pool %>% tidy_output()
```

## Adjusted model 3 frequency dissatisfaction

* All Model 1 variables   
* All Model 2 variables   
* Night sweats   
* Hot flashes   
* Vaginal dryness   
* Current use of female hormone therapy 

```{r}
m3_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse4cat_f + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f + night_sweats_f + hot_flashes_f + vag_dry_f +
      hormnw_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m3_freq_dissat_pool <- pool(m3_freq_dissat)
m3_freq_dissat_pool %>% tidy_output()
```
