---
title: "Preprocess 02: Preliminary Variable Management"
date: "`r Sys.Date()`"
output: 
  github_document: default
---

```{r load_packages, message=FALSE}
# Load packages
library(tidyverse)
library(feather)
library(data.table)

# Load functions
source("../R scripts/functions.R")
```

-------------------------------------------------------------------------------

# Dichotomize medications to reduce duplicated days

-------------------------------------------------------------------------------

* Tccode creates a separate row within id and day for each medication taken. We only care about SSRI's. We will collapse into a single row within id and day with a single variable indicating if an SSRI was reported or not.

* Drug class from Erika (2017-03-09): 581600

```{r load_analysis_01}
# Load data
analysis_01 <- read_feather("../data/analysis_01.feather")
check_data(analysis_01) # 2,448,638 observations and 52 variables
```

```{r}
dt <- as.data.table(analysis_01)
```

## Tag observations with 581600 and distribute SSRI within id and days

```{r tag_ssri, rows.print=12}
# Create new var, SSRI, that is equal to 1 if any drug within id and day is an SSRI
dt[, ssri := if_else(any(tccode == "581600"), 1, 0, NA_real_), by = .(id, days)]
# select(dt, id, days, tccode, ssri) %>% filter(id == 100088)
check_data(dt) # 2,448,638 observations and 53 variables
```

## Now that SSRI is distributed within id and day, keep one row per day 

```{r one_row_per_ssri}
# Drop tccode
dt[, tccode := NULL]
check_data(dt) # 2,448,638 observations and 52 variables
```

```{r check_dups}
# Create new logical var, dup, that is TRUE when the row is a duplicate
dt[, dup := duplicated(dt)]
check_data(dt) # 2,448,638 observations and 53 variables
```

```{r}
# Count dup == TRUE
sum(dt$dup) # 1,016,190
```

```{r drop_dups}
# Drop duplicate rows
dt <- dt[dup == FALSE, ]
check_data(dt) # 1,432,448 observations and 53 variables
```

```{r check_dups_02}
# Any duplicates by id and days?
dt[, dup := duplicated(.SD), .SDcols = c("id", "days")]
dt[, sum(dup)] # 0
```

```{r drop_dups_02}
# Drop dup
dt$dup <- NULL
check_data(dt) # 1,432,448 observations and 52 variables
```

```{r save_analysis_02}
# Save progress
analysis_02 <- dt
write_feather(analysis_02, path = "../data/analysis_02.feather")
```

At this point there are 1,432,448 observations and 52 variables. There are no duplicate rows in terms of id and days.









-------------------------------------------------------------------------------

# Expand age across time

-------------------------------------------------------------------------------

```{r load_analysis_02}
# Load data
analysis_02 <- read_feather("../data/analysis_02.feather")
check_data(analysis_02) # 1,432,448 observations and 52 variables
```

```{r}
dt <- as.data.table(analysis_02)
```

```{r wrangle_age}
dt[, age_days := age * 365.25]
dt[, age_days_2 := na.omit(age_days)[1], by = id]
dt[, age_days_2 := age_days_2 + days]
dt[, age_2 := round(age_days_2 / 365.25)]
# select(dt, id, days, age, age_days, age_days_2, age_2)
dt[, age := age_2]
dt[, c("age_days", "age_days_2", "age_2") := NULL]
# select(dt, id, days, age)
dt[days == 0, .(.N, `Mean Age at Baseline` = mean(age), SD = sd(age), Min = min(age), Max = max(age))]
```

```{r save_analysis_03}
# Save progress
analysis_03 <- dt
write_feather(analysis_03, path = "../data/analysis_03.feather")
```









-------------------------------------------------------------------------------

# Create variables: obs, finalobs, finalage, numobs, finalyears

-------------------------------------------------------------------------------

```{r load_analysis_03}
# Load data
analysis_03 <- read_feather("../data/analysis_03.feather")
check_data(analysis_03) # 1,432,448 observations and 52 variables
```

```{r}
dt <- as.data.table(analysis_03)
```

## Creating a observation number variable

```{r create_obs}
dt <- dt[order(id, days), ]
dt[, obs := 1:.N, by = id]
# select(dt, id, days, obs)
dt[, .(`Mean Number of Observations Per Woman` = mean(obs), SD = sd(obs))]
```
			
## Creating an age at final observation tag

```{r tag_age_final_obs}
dt[, final_obs := if_else(obs == max(obs), TRUE, FALSE), by = id]
attributes(dt$final_obs)$label <- "Tag largest observation number by id"
# select(dt, id, days, obs, final_obs)
sum(dt$final_obs) # 161,808
```

## Creating age at final observation variable

```{r create_age_final_obs}
dt[final_obs == TRUE, final_age := age, by = id]
dt[final_obs == TRUE, .(.N, `Mean Age at Last Observation` = mean(final_age), SD = sd(final_age), 
                        Min = min(final_age), Max = max(final_age))]
```

## Distribute final age across observations

```{r dist_final_age}
dt[, final_age := max(final_age, na.rm = TRUE), by = id]
attributes(dt$final_age)$label <- "Age at final observation"
# select(dt, id, days, age, final_age)
```

## Creating number of observations variable at final observation

```{r create_n_obs_at_final_obs}
dt[final_obs == TRUE, num_obs := obs]
dt[final_obs == TRUE, .(.N, `Mean Observations per Woman` = mean(obs), SD = sd(obs), Min = min(obs), 
                        Max = max(obs))]
```

## Distribute num_obs across observations

```{r dist_num_obs}
dt[, num_obs := max(num_obs, na.rm = TRUE), by = id]
attributes(dt$num_obs)$label <- "Number of observations"
# select(dt, id, days, obs, num_obs)
```

## Creating years since e/r at final observation variable

```{r create_years_since_enroll}
dt[, years := round(days / 365.25)]
dt[final_obs == TRUE, total_years := years]
dt[final_obs == TRUE, .(.N, `Mean Number of Years Enrolled` = mean(years), SD = sd(years), Min = min(years), 
                        Max = max(years))]
```

## Distribute total_years across observations

```{r dist_years_since_enroll}
dt[, total_years := max(total_years, na.rm = TRUE), by = id]
attributes(dt$total_years)$label <- "Total years since e/r @ last observation"
# select(dt, id, days, final_obs, years, total_years)
```

```{r save_analysis_04}
# Save progress
analysis_04 <- dt
write_feather(analysis_04, path = "../data/analysis_04.feather")
```

```{r session_info, echo=FALSE}
sessionInfo()
```