---
title: 'Preprocess 01: Read in SAS Data Set'
date: '`r Sys.Date()`'
output:
  github_document: default
---

---

* Individual form data sets were previously cleaned and duplicates (by id, days) were removed using SAS. However, after merging forms, there may be new duplicate rows by id/days.

* Individual form data sets were previously merged by id and days in the program merge_forms.

```{r load_packages, message=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(haven)
library(data.table)

# Load functions
source("../R scripts/functions.R")
```

```{r}
# Load data created in SAS
start  <- now()
merged <- read_sas("../data/merged.sas7bdat")
finish <- now()
difftime(finish, start)
```

```{r}
check_data(merged) # 2,448,638 observations and 815 variables
```

Quick Save

```{r}
write_rds(merged, path = "../data/merged.rds")
```

Load data again (if needed)

```{r}
# merged <- read_rds("../data/merged.rds")
# check_data(merged) # 2,448,638 observations and 815 variables
```

# Set all variable names to lowercase

```{r}
names(merged) <- tolower(names(merged))
```

# Subset variables of interest

```{r}
analysis_01 <- merged %>% 
  select(
    # Administrative and Sociodemographic
	  id, days, age, ager, race_eth, edu4cat, inc5cat, inc5cat_f20, marital, married, sex, ctos, parity,

	  # Health Behavior
	  texpwk, alcswk, f60alc, f60alcwk, f60caff, smoknow, smoking, horm, hormnw, tccode, livalor, livaln,

	  # Health and Wellness
	  lifequal, pshtdep, bmi, genhel, hyst, nightswt, hotflash, vagdry, incont, atrophy,

	  # Chronic Disease
	  arthrit, brca_f30, cervca, endo_f30, ovryca, cvd, diab, hypt, osteopor, pad,

	  # Sexual Function
	  sexactiv, satsex, satfrqsx,

	  # Abuse
	  phyab, verbab
  )
check_data(analysis_01) # 2,448,638 observations and 50 variables
```


# Change " " in character vectors to NA

```{r}
analysis_01 <- analysis_01 %>% 
  map_if(is.character, function(x) {
    x[x == ""] <- NA
    x
  }) %>% 
  as.tibble()
```

# Data checks

```{r}
check_data(analysis_01) # 2,448,638 observations and 50 variables
```

```{r}
count_ids(analysis_01$id) # 161,808 unique women
```

```{r}
sum(analysis_01$days == 0) # There are 166,347 day 0's, but only 161,808 women.
```

Why do some women have more than one day 0?

```{r}
check <- analysis_01 %>%
  group_by(id) %>%
  mutate(
    day0 = if_else(days == 0, 1L, NA_integer_),
    count_day0 = sum(day0, na.rm = TRUE)
  ) %>%
  filter(count_day0 > 1)
```

Manually review. The multiple days come from form 44 - medications. Each separate medication has a separate row.

```{r}
rm(check, start, finish)
```

Save in R binary format

```{r}
write_rds(analysis_01, path = "../data/analysis_01.rds")
```

```{r session_info, echo=FALSE}
sessionInfo()
```
