---
title: "Table 2. Unadjusted and adjusted odds of global sexual dissatisfaction and sexual frequency dissatisfaction at baseline, results from the Women’s Health Initiative (WHI) (N = 83,218)."
output:
  word_document:
    reference_docx: word_style_template_01.docx
---

```{r load_packages,include=FALSE}
# Load packages
library(tidyverse)
library(mice)
library(forcats)
library(broom)
library(modelr)
```

```{r load_data, include=FALSE}
# Load data
imp <- read_rds(path = "../data/imp.rds")
```

```{r create_table_header, include=FALSE}
table_02 <- tibble(
  `Recent Abuse` = vector(mode = "character", length = 3),
  `Unsatisfied with Sexual Activity Unadjusted` = vector(mode = "character", length = 3),
  `Unsatisfied with Sexual Activity Adjusted` = vector(mode = "character", length = 3),
  `Unsatisfied with Sexual Activity Frequency Unadjusted` = vector(mode = "character", length = 3),
  `Unsatisfied with Sexual Activity Frequency Adjusted` = vector(mode = "character", length = 3)
)
table_02$`Recent Abuse` <- c("", "No", "Yes")
table_02[1, 2:5] <- "OR (95% CI)"
table_02[2, 2:5] <- "Reference"
```

```{r create_table_02, include=FALSE}
# Unsatisfied with Sexual Activity Unadjusted
m_mi <- with(imp,
  glm(
    fct_rev(satisfied_f) ~ abuse_d_f,
    family = binomial(link = "logit")
  )
)
mi_est       <- pool(m_mi)
mi_est       <- summary(mi_est) %>% tidy()
mi_est$OR    <- exp(mi_est$est) %>% round(2)
mi_est$lci   <- exp(mi_est$lo.95) %>% round(2)
mi_est$uci   <- exp(mi_est$hi.95) %>% round(2)
mi_est$or_95 <- paste0(mi_est$OR, " (", mi_est$lci, " - ", mi_est$uci, ")")
mi_est       <- select(mi_est, .rownames, or_95)

table_02[3, 2] <- mi_est[2, 2]

# Unsatisfied with Sexual Activity Adjusted
m_mi <- with(imp,
  glm(
    fct_rev(satisfied_f) ~ abuse_d_f + age + race_eth_f + edu4cat_f + 
      inc5cat_f + married_f + sex_f + ctos_f + parity_f + texpwk + alcswk + 
      f60caff + smoking_f + horm_f + hormnw_f + ssri_f + lifequal + pshtdep + 
      bmi + good_health_f + hyst_f + night_sweats_f + hot_flashes_f + 
      vag_dry_f + incont_f + chronic_disease_f,
    family = binomial(link = "logit")
  )
)
mi_est       <- pool(m_mi)
mi_est       <- summary(mi_est) %>% tidy()
mi_est$OR    <- exp(mi_est$est) %>% round(2)
mi_est$lci   <- exp(mi_est$lo.95) %>% round(2)
mi_est$uci   <- exp(mi_est$hi.95) %>% round(2)
mi_est$or_95 <- paste0(mi_est$OR, " (", mi_est$lci, " - ", mi_est$uci, ")")
mi_est       <- select(mi_est, .rownames, or_95)

table_02[3, 3] <- mi_est[2, 2]

# Unsatisfied with Sexual Activity Frequency Unadjusted
m_mi <- with(imp,
  glm(
    fct_rev(freq_satisfied_f) ~ abuse_d_f,
    family = binomial(link = "logit")
  )
)
mi_est       <- pool(m_mi)
mi_est       <- summary(mi_est) %>% tidy()
mi_est$OR    <- exp(mi_est$est) %>% round(2)
mi_est$lci   <- exp(mi_est$lo.95) %>% round(2)
mi_est$uci   <- exp(mi_est$hi.95) %>% round(2)
mi_est$or_95 <- paste0(mi_est$OR, " (", mi_est$lci, " - ", mi_est$uci, ")")
mi_est       <- select(mi_est, .rownames, or_95)

table_02[3, 4] <- mi_est[2, 2]

# Unsatisfied with Sexual Activity Frequency Adjusted
m_mi <- with(imp,
  glm(
    fct_rev(freq_satisfied_f) ~ abuse_d_f + age + race_eth_f + edu4cat_f + 
      inc5cat_f + married_f + sex_f + ctos_f + parity_f + texpwk + alcswk + 
      f60caff + smoking_f + horm_f + hormnw_f + ssri_f + lifequal + pshtdep + 
      bmi + good_health_f + hyst_f + night_sweats_f + hot_flashes_f + 
      vag_dry_f + incont_f + chronic_disease_f,
    family = binomial(link = "logit")
  )
)
mi_est       <- pool(m_mi)
mi_est       <- summary(mi_est) %>% tidy()
mi_est$OR    <- exp(mi_est$est) %>% round(2)
mi_est$lci   <- exp(mi_est$lo.95) %>% round(2)
mi_est$uci   <- exp(mi_est$hi.95) %>% round(2)
mi_est$or_95 <- paste0(mi_est$OR, " (", mi_est$lci, " - ", mi_est$uci, ")")
mi_est       <- select(mi_est, .rownames, or_95)

table_02[3, 5] <- mi_est[2, 2]
```

```{r, echo=FALSE}
table_02_kable <- knitr::kable(table_02)
write_rds(table_02_kable, path = "table_02_kable.rds") # Save for use in dashboard
table_02_kable
```

1. Sexual satisfaction measured as level of satisfaction with current sexual activities, either with a partner or alone. 
2. Sexual frequency satisfaction measured as satisfaction with the frequency of sexual activity.
3. Adjusted for relevant health history (including depression) and sociodemographic characteristics.
4. Recent abuse measured as physical abuse and/or verbal abuse experienced within one year of baseline.

```{r notes, include=FALSE}
# After knitting word table:
# Change all font to TNR 11
# Left justify title
# Remove bold from title - except "Table 2"
# Center columns 3 and above from top to bottom
# Add bottom border to table
# Set footnotes to 10 pt font
```


