---
title: "Logistic models for Table 2"
date: "Created: 2017-05-29 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    css: custom-css.css
---

# Overview

In this file, we start with the multiply imputed baseline data for all women who were sexually active in the previous year. We then use logistic regression to model the relationship between abuse and sexual satisfaction.

From EK 2017-07-17: 

**Compute odds of sexual satisfaction (Table 2).**

Conduct adjusted models to examine cross-sectional odds of sexual satisfaction (Global/frequency), with sequential modeling as described below. Use multiple imputation to account for the missingness of chronic disease information. Sequential modeling as follows:

**Model 1 (Sociodemographics):**   

* Age   
* Race & Ethnicity   
* Education   
* Income   
* Marital Status   
* WHI study component   
	
**Model 2 (Health/health risk factors + Model 1):**   

* All Model 1 variables   
* Pack years of smoking   
* Body mass index (BMI)   
* Hysterectomy ever   
* Urinary incontinence ever   
* Hypertension ever (variable label = hypt; self-report Form 30)   
* Cardiovascular disease ever (variable label = CVD; self-report Form 30)   
* Arthritis ever (Label = arthrit; self-report Form 30)   
* Cancer ever (Label = canc_f30; Form 30)   
* Diabetes ever (Variable label = diab & diabtrt; self-report Form 2)   
* Hip fracture age 55 or older (Variable label = hip55; self-report Form 30)   
* Self-rated health   
* Depressive symptoms   
* Current use of SSRI   

**Model 3 (Menopausal symptoms/treatment variables that impact menopausal + Model 1 +Model 2):**  

* All Model 1 variables   
* All Model 2 variables   
* Night sweats   
* Hot flashes   
* Vaginal dryness   
* Current use of female hormone therapy   

# Table of contents

1. [Exploratory data analysis](#eda)
2. [Unadjusted model dissatisfaction](#unadjusted-sat)
3. [Unadjusted model frequency dissatisfaction](#unadjusted-freq-sat)

```{r load_packages,include=FALSE}
Sys.setenv(TZ = "US/Central")
library(tidyverse)
library(feather)
library(bfuncs)
library(mice)
library(broom)
library(modelr)
```

Create helper function for displaying model output

```{r}
tidy_output <- function(x, ...) {
  x %>% 
    summary() %>% 
    tidy() %>% 
    mutate(
    est = exp(est) %>% round(2) %>% format(nsmall = 2L),
    lo.95 = exp(lo.95) %>% round(2) %>% format(nsmall = 2L),
    hi.95 = exp(hi.95) %>% round(2) %>% format(nsmall = 2L),
    or_95 = paste0(est, " (", lo.95, " - ", hi.95, ")")
  ) %>% 
  select(.rownames, est, lo.95, hi.95, or_95)
}
```

```{r load_data, include=FALSE}
mids_83329 <- read_rds(path = "../data/mids_83329.rds")
analysis_11 <- read_feather("../data/analysis_11.feather")
```










-------------------------------------------------------------------------------

# Exploratory data analysis {#eda}

-------------------------------------------------------------------------------

```{r}
about_data(analysis_11) # 916,619 observations and 59 variables
```

What does the distribution of satisfied by abuse look like?

```{r}
analysis_11 %>% 
  filter(!.imp == 0) %>% 
  group_by(abuse, satisfied_f) %>% 
  freq_table() %>% 
  filter(col_cat == "Yes") # satisfied_f
```

```{r}
analysis_11 %>% 
  filter(!.imp == 0) %>%
  group_by(abuse, freq_satisfied_f) %>% 
  freq_table() %>% 
  filter(col_cat == "Yes") # freq_satisfied_f
```










-------------------------------------------------------------------------------

# Unadjusted model dissatisfaction {#unadjusted-sat}

-------------------------------------------------------------------------------

_Using mice::with will return a Multiply imputed repeated analyses (MIRA) object_

```{r}
unadj_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse,
    family = binomial(link = "logit")
  )
)
```

```{r}
unadj_dissat_pool <- pool(unadj_dissat)
unadj_dissat_pool %>% tidy_output()
```


## Adjusted model 1 dissatisfaction

* Age   
* Race & Ethnicity   
* Education   
* Income   
* Marital Status   
* WHI study component   

```{r}
m1_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m1_dissat_pool <- pool(m1_dissat)
m1_dissat_pool %>% tidy_output()
```


## Adjusted model 2 dissatisfaction

* All Model 1 variables   
* Pack years of smoking   
* Body mass index (BMI)   
* Hysterectomy ever   
* Urinary incontinence ever   
* Hypertension ever (variable label = hypt; self-report Form 30)   
* Cardiovascular disease ever (variable label = CVD; self-report Form 30)   
* Arthritis ever (Label = arthrit; self-report Form 30)   
* Cancer ever (Label = canc_f30; Form 30)   
* Diabetes ever (Variable label = diab & diabtrt; self-report Form 2)   
* Hip fracture age 55 or older (Variable label = hip55; self-report Form 30)   
* Self-rated health   
* Depressive symptoms   
* Current use of SSRI  

```{r}
m2_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m2_dissat_pool <- pool(m2_dissat)
m2_dissat_pool %>% tidy_output()
```


## Adjusted model 3 dissatisfaction

* All Model 1 variables   
* All Model 2 variables   
* Night sweats   
* Hot flashes   
* Vaginal dryness   
* Current use of female hormone therapy 

```{r}
m3_dissat <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f + night_sweats_f + hot_flashes_f + vag_dry_f +
      hormnw_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m3_dissat_pool <- pool(m3_dissat)
m3_dissat_pool %>% tidy_output()
```

```{r}
rm(m1_dissat, m1_dissat_pool, m2_dissat, m2_dissat_pool, m3_dissat, 
   m3_dissat_pool, unadj_dissat, unadj_dissat_pool)
```










-------------------------------------------------------------------------------

# Unadjusted model frequency dissatisfaction {#unadjusted-freq-sat}

-------------------------------------------------------------------------------

```{r}
unadj_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse,
    family = binomial(link = "logit")
  )
)
```

```{r}
unadj_freq_dissat_pool <- pool(unadj_freq_dissat)
unadj_freq_dissat_pool %>% tidy_output()
```


## Adjusted model 1 frequency dissatisfaction

* Age   
* Race & Ethnicity   
* Education   
* Income   
* Marital Status   
* WHI study component   

```{r}
m1_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m1_freq_dissat_pool <- pool(m1_freq_dissat)
m1_freq_dissat_pool %>% tidy_output()
```


## Adjusted model 2 frequency dissatisfaction

* All Model 1 variables   
* Pack years of smoking   
* Body mass index (BMI)   
* Hysterectomy ever   
* Urinary incontinence ever   
* Hypertension ever (variable label = hypt; self-report Form 30)   
* Cardiovascular disease ever (variable label = CVD; self-report Form 30)   
* Arthritis ever (Label = arthrit; self-report Form 30)   
* Cancer ever (Label = canc_f30; Form 30)   
* Diabetes ever (Variable label = diab & diabtrt; self-report Form 2)   
* Hip fracture age 55 or older (Variable label = hip55; self-report Form 30)   
* Self-rated health   
* Depressive symptoms   
* Current use of SSRI  

```{r}
m2_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m2_freq_dissat_pool <- pool(m2_freq_dissat)
m2_freq_dissat_pool %>% tidy_output()
```


## Adjusted model 3 frequency dissatisfaction

* All Model 1 variables   
* All Model 2 variables   
* Night sweats   
* Hot flashes   
* Vaginal dryness   
* Current use of female hormone therapy 

```{r}
m3_freq_dissat <- with(mids_83329, 
  glm(
    freq_dissatisfied ~ abuse + age + race_eth_f + edu2cat_f + inc5cat_f + 
      married_f + ctos_f + packyrs + bmi + hyst_f + incont_f + hypt_f +
      cvd_f + arthrit_f + canc_f + diab_f + hip55_f + good_health_f + 
      pshtdep + ssri_f + night_sweats_f + hot_flashes_f + vag_dry_f +
      hormnw_f,
    family = binomial(link = "logit")
  )
)
```

```{r}
m3_freq_dissat_pool <- pool(m3_freq_dissat)
m3_freq_dissat_pool %>% tidy_output()
```

```{r}
rm(m1_freq_dissat, m1_freq_dissat_pool, m2_freq_dissat, m2_freq_dissat_pool,
   m3_freq_dissat, m3_freq_dissat_pool, unadj_freq_dissat, 
   unadj_freq_dissat_pool)
```

&nbsp;

-------------------------------------------------------------------------------

```{r}
sessionInfo()
```