---
title: "Check for age by abuse interaction"
date: "Created: 2018-01-15 <br> Updated: `r Sys.Date()`"
output: html_notebook
---

<style>
  hr {
    background-color: #66b3ff;
    height: 1px;
  }
</style>

# Overview

In this file, we start with the multiply imputed baseline data for all women who were sexually active in the previous year. We are checking to see if there is the effect of abuse differs by age.

Check:    
1. Raw data - no interaction   
2. Raw data - interaction - age*abuse   
3. Raw data - interaction - age_c*abuse    
4. Mids_83329 - no interaction   
5. Mids_83329 - interaction   
6. Mids_interaction - no interaction   
7. Mids_interaction - interaction - age*abuse    
8. Mids_interaction - interaction - age_c*abuse    

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r load_packages, message=FALSE}
library(tidyverse)
library(feather)
library(bfuncs)
library(mice)
library(broom)
library(modelr)
```

```{r load_data, include=FALSE}
baseline_only <- read_feather("../data/baseline_only.feather")
mids_interaction <- read_rds(path = "../data/mids_interaction.rds")
mids_83329 <- read_rds(path = "../data/mids_83329.rds")
analysis_11_interaction <- read_feather("../data/analysis_11_interaction.feather")
```

Data management

```{r}
baseline_only <- baseline_only %>% 
  mutate(
    abuse = case_when(
      abuse4cat_f == "Did not experience abuse" ~ 0,
      abuse4cat_f == "Verbal abuse only" ~ 1,
      abuse4cat_f == "Physical abuse only" ~ 1,
      abuse4cat_f == "Physical and verbal abuse" ~ 1
    ),
    satisfied = if_else(satisfied_f == "Yes", 1L, 0L, NA_integer_),
    dissatisfied = if_else(satisfied == 1, 0L, 1L, NA_integer_),
    age_c = age - 60
  )
```

Helper functions

```{r}
tidy_glm <- function(x, ...) {
 out <- tidy(x)
 ci  <- confint(x) %>% exp() %>% as.data.frame()
 out <- bind_cols(out, ci)
 out <- out %>%
   mutate(
     estimate = exp(estimate),
     p_value = round(p.value, 10) %>% format(nsmall = 6L)
   ) %>% 
   select(-p.value)
 out
}

tidy_output_mids <- function(x, ...) {
  x %>% 
    summary() %>% 
    tidy() %>% 
    mutate(
    est = exp(est) %>% round(2) %>% format(nsmall = 2L),
    lo.95 = exp(lo.95) %>% round(2) %>% format(nsmall = 2L),
    hi.95 = exp(hi.95) %>% round(2) %>% format(nsmall = 2L),
    p_value = round(`Pr...t..`, 10) %>% format(nsmall = 6L),
    or_95 = paste0(est, " (", lo.95, " - ", hi.95, ")")
  ) %>% 
  select(.rownames, est, lo.95, hi.95, or_95, p_value)
}
```

# 1. Raw data - no interaction

```{r message=FALSE}
glm(dissatisfied ~ abuse + age, data = baseline_only) %>% 
  tidy_glm()
```

# 2. Raw data - interaction - age*abuse   

```{r message=FALSE}
glm(dissatisfied ~ abuse*age, data = baseline_only) %>% 
  tidy_glm()
```

# 3. Raw data - interaction - age_c*abuse  

```{r message=FALSE}
glm(dissatisfied ~ abuse*age_c, data = baseline_only) %>% 
  tidy_glm()
```

# 4. Mids_83329 - no interaction   

```{r}
m4 <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse + age,
    family = binomial(link = "logit")
  )
)
```

```{r}
m4_pool <- pool(m4)
m4_pool %>% tidy_output_mids()
```

# 5. Mids_83329 - interaction   

```{r}
m5 <- with(mids_83329, 
  glm(
    dissatisfied ~ abuse*age,
    family = binomial(link = "logit")
  )
)
```

```{r}
m5_pool <- pool(m5)
m5_pool %>% tidy_output_mids()
```

# 6. Mids_interaction - no interaction 

```{r}
m6 <- with(mids_interaction, 
  glm(
    dissatisfied ~ abuse + age,
    family = binomial(link = "logit")
  )
)
```

```{r}
m6_pool <- pool(m6)
m6_pool %>% tidy_output_mids()
```

# 7. Mids_interaction - interaction - age*abuse  

```{r}
m7 <- with(mids_interaction, 
  glm(
    dissatisfied ~ abuse*age,
    family = binomial(link = "logit")
  )
)
```

```{r}
m7_pool <- pool(m7)
m7_pool %>% tidy_output_mids()
```

# 8. Mids_interaction - interaction - age_c*abuse 

```{r}
stacked <- complete(mids_interaction, action = "long", include = TRUE)
```

```{r}
stacked <- stacked %>% 
  mutate(age_c = age - 60)
```

```{r}
mids_interaction_age_c <- as.mids(stacked)
```

```{r}
m8 <- with(mids_interaction_age_c, 
  glm(
    dissatisfied ~ abuse*age_c,
    family = binomial(link = "logit")
  )
)
```

```{r}
m8_pool <- pool(m8)
m8_pool %>% tidy_output_mids()
```

```{r echo=FALSE}
sessionInfo()
```
