---
title: "Preprocess 05: Clean and Manage Covariates"
date: "Created: 2017-07-17 <br> Updated: `r Sys.Date()`"
output: html_notebook
---

<style>
  hr {
    background-color: #66b3ff;
    height: 1px;
  }
</style>

# Overview

In this file, we do general preprocessing of covariate data. Specifically, we:

* Recode missing   
* Create factor variables   
* Distribute time-invariant variables across all observations (e.g. race)   
* Distribute "have you ever" variables (e.g. diabetes ever)   
* Carry forward relatively time-stable variables (e.g. income)   
* Perform variable transformations

# Table of contents

1. [Distributing selected covariates across observations](#distribute)   
2. [Sociodemographic covariates](#socio)   
3. [Health and health risk factors](#health)   
4. [Menopausal symptoms and treatment variables](#menopausal)   
5. [Save progress](#save)   

```{r load_packages, message=FALSE}
Sys.setenv(TZ = "US/Central")
library(feather)
library(tidyverse)
library(data.table)
library(zoo)
library(gmodels)
library(bfuncs)
source("../r/functions.R")
```

```{r load_analysis_07}
analysis_07 <- read_feather("../data/analysis_07.feather")
about_data(analysis_07) # 1,432,448 observations and 93 variables
```

```{r}
dt <- as.data.table(analysis_07)
```








-------------------------------------------------------------------------------

# Distributing selected covariates across observations {#distribute}

-------------------------------------------------------------------------------

```{r time_invariant}
# Variables that are either time-invariant, or considered to be nearly time-invariant in this sample
time_invariant <- c("ager", "race", "ethnic", "edu4cat","sex", "ctos", "parity")

# Distribute time-invariant variables across all observations within id
dt[, (time_invariant) := lapply(.SD, function(x) na.omit(x)[1]), 
   .SDcols = time_invariant, 
   by = id]
```

```{r check_ever_vars}
# Check "have you ever" variables
ever_vars <- c("horm", "hyst", "arthrit", "brca_f30", "cervca", "endo_f30", "ovryca", "cvd", "diab", 
               "hypt", "osteopor", "pad", "canc_f30", "hip55", "diabtrt", "incont")
dt[, c("days", ever_vars), with = FALSE]
```

```{r ever_vars}
# Distribute "have you ever" variables
# If they say "no", then all previous observations should be no. If you've never had diabetes on day 100, then you didn't have diabetes on day 99.
# If they say "Yes", then all future observations should be yes. If you have diabetes on day 100, then you have ever had diabetes on day 200.
dt[, (ever_vars) := lapply(.SD, distribute_ever), 
   .SDcols = ever_vars, 
   by = id]
```

```{r check_ever_vars_again}
dt[, c("id", "days", ever_vars), with = FALSE]
```

How many logic errors (0's after 1's) were there?

```{r rows.print=16}
nines <- function(x) {
  i   <- which(x == 9)
  ids <- dt$id[i]
  ids <- unique(ids)
  out <- length(ids)
  out
}

dt[, lapply(.SD, nines), .SDcols = ever_vars] %>% 
  gather("Variable", "N ID's with Logic Error")

rm(nines)
```

There are 15,859 people with a 9 for incont. Nines appear when there are logically inconsistent answers to an "ever" question. For example, they say on day 0 that they have ever leaked urine and then say no to the same question on day 100.

```{r}
# nines <- dt %>% filter(incont == 9) %>% pull(id) %>% unique()
# analysis_07 %>% filter(id %in% nines) %>% select(id, days, incont)
```

There are a lot of these logical inconsistencies. They are typically at the very last observation. We will handle this by making every observation after the first "1" a "1".

```{r}
dt <- dt %>% 
  mutate(incont = if_else(incont == 9, 1, incont))
```

```{r}
# Save progress
analysis_08 <- dt
write_feather(analysis_08, path = "../data/analysis_08.feather")
```








-------------------------------------------------------------------------------

# Sociodemographic covariates {#socio}

-------------------------------------------------------------------------------

```{r}
# Load data
analysis_08 <- read_feather("../data/analysis_08.feather")
about_data(analysis_08) # 1,432,448 observations and 93 variables
```

```{r}
dt <- as.data.table(analysis_08)
```

### Age

[Age at screening](https://www.whi.org/researchers/data/WhiDataDict/dem_ctos_inv.pdf)

```{r}
dt[, .(Missing = sum(is.na(age)), Mean = mean(age), SD = sd(age))] # Over all observations
```

### Age group

[Age group at screening](https://www.whi.org/researchers/data/WhiDataDict/dem_ctos_inv.pdf)

```{r}
# Rename ager
setnames(dt, "ager", "age_group")

# Create factor
dt[, age_group_f := factor(age_group, labels = c("<50-59", "60-69", "70-79+"))]

# Descriptives at first obs
dt[obs == 1, .(Women = .N), by = age_group_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Race / Ethnicity

[Ethnicity](https://www.whi.org/researchers/data/WhiDataDict/dem_ctos_inv.pdf)

```{r}
# Recode ethnic
dt <- dt %>% mutate(race_eth = case_when(
    is.na(ethnic) ~ NA_real_, # Missing
    ethnic == 5   ~ 1,        # White (not of Hispanic origin)
    ethnic == 3   ~ 2,        # Black or African-American
    ethnic == 4   ~ 3,        # Hispanic/Latino
    ethnic == 2   ~ 4,        # Asian or Pacific Islander
    ethnic == 1   ~ 5,        # American Indian or Alaskan Native
    TRUE          ~ 6         # Other
  )
)
dt <- as.data.table(dt)
```

```{r}
# Create factor
dt[, race_eth_f := factor(race_eth, levels = c(1:6), 
                          labels = c("White", "Black or African American", 
                                     "Hispanic/Latina", "Asian/Pacific Islander", 
                                     "American Indian/Alaska Native", "Other"))]

# Descriptives at first obs
dt[obs == 1][order(race_eth_f), .(Women = .N), by = race_eth_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
# dt %>% filter(is.na(race_eth)) %>% select(id, days, race_eth) %>% summarise(length(unique(id)))
# 413 people matches WHI documentation
```

### Education

[Edcuation](https://www.whi.org/researchers/data/WhiDataDict/dem_ctos_inv.pdf)

```{r}
# Create factor
dt[, edu4cat_f := factor(edu4cat, labels = c("<HS", "HS Grad", "Some College or Tech School", "College Grad"))]

# Descriptives at first obs
dt[obs == 1][order(edu4cat_f), .(Women = .N), by = edu4cat_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
# 1,216 missing matches WHI documentation
```

```{r}
# Dichotomize Education
dt[edu4cat_f %in% c("<HS", "HS Grad", "Some College or Tech School"), edu2cat := 0]
dt[edu4cat_f == "College Grad", edu2cat := 1]
dt[, edu2cat_f := factor(edu2cat, labels = c("Not College Graduate", "College Graduate"))]
dt[obs == 1][order(edu2cat_f), .(Women = .N), by = edu2cat_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Income

[Family income](https://www.whi.org/researchers/data/WhiDataDict/dem_ctos_inv.pdf)

```{r}
# How stable is income over time?

# Take a look
# select(dt, id, days, inc5cat, inc5cat_f20)

# How often are inc5cat and inc5cat_f20 measured on the same day?
# 3,148
# On those 3,148 days, how often do inc5cat and inc5cat_f20 disagree?
# dt[!is.na(inc5cat) & !is.na(inc5cat_f20), .(id, days, inc5cat, inc5cat_f20)][inc5cat != inc5cat_f20]
# All on day zero and all agree

# Create a combined income variable
dt[!is.na(inc5cat) & !is.na(inc5cat_f20), combined := inc5cat]
dt[is.na(combined), combined := sum(inc5cat, inc5cat_f20, na.rm = TRUE), by = .(id, days)]
dt[combined == 0, combined := NA_real_]
# select(dt, id, days, inc5cat, inc5cat_f20, combined)

# Does combined change from measure to measure?
dt[!is.na(combined), length(unique(combined)) != 1, by = id][
  , .(`Women with Income` = .N, `Income change` = sum(V1))]
```

Income is time-invariant. Will distribute across all observations.

```{r}
# Drop unneeded income variables
dt[, c("inc5cat_f20", "combined") := NULL]

# Distribute across observations
dt[, inc5cat := na.omit(inc5cat)[1], by = id]

# Create factor
dt[, inc5cat_f := factor(inc5cat, labels = c("20,000 or less", "20-34,999", "35-49,999", "50-74,999",
                                             "75,000+"))]

# Descriptives at first obs
dt[days == 0][order(inc5cat_f), .(Women = .N), by = inc5cat_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
# 10,874 missing match WHI documentation for missing and don't know.
```

Add two-level for Table 1

```{r}
dt[inc5cat_f %in% c("20,000 or less", "20-34,999", "35-49,999"), inc2cat := 1]
dt[inc5cat_f %in% c("50-74,999", "75,000+"), inc2cat := 2]
dt[, inc2cat_f := factor(inc2cat, labels = c("< $50,000", "$50,000+"))]

dt[days == 0][order(inc2cat_f), .(Women = .N), by = inc2cat_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Marital status

* [Married](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf) is worded, "Are you currently married or in an intimate relationship with at least one person?"

* [Marital](https://www.whi.org/researchers/data/WhiDataDict/f20_ctos_inv.pdf) is worded, "What is your current marital status? (Mark the one that best describes you.)":   
1. Never married    
2. Divorced or separated    
3. Widowed   
4. Presently married   
5. Marriage-like relationship    

Recode the marital to 1, 2, and 3 vs. 4 & 5. Should make married and marital roughly comparible.

```{r}
dt[, marital := if_else(marital %in% c(1, 2, 3), 0, 
                if_else(marital %in% c(4, 5),    1,
                NA_real_))]
# select(dt, id, days, married, marital)
```

```{r}
# Carry forward married
dt[, married := na.locf(married, na.rm = FALSE), by = id]

# Carry forward marital
dt[, marital := na.locf(marital, na.rm = FALSE), by = id]
# select(dt, id, days, married, marital)
```

```{r}
# Which variable has the greatest amount of missing?

cat("Number of missing values for married =", format(sum(is.na(dt$married)), big.mark = ",")) # 373,512
cat("\n")
cat("Number of missing values for marital =", format(sum(is.na(dt$marital)), big.mark = ",")) # 263,161
```

Marital has fewer missing values. 

We previously tried augmenting missing values in marital with data from married - when available. However, sometime there were disagreements. Trying to pick a winner in the disagreements may introduce bias. Setting disagreements to NA results in a greater number of missing than just using marital as it is. So, we'll just use marital as it is.

```{r}
# Drop married and rename marital
dt$married <- NULL
setnames(dt, "marital", "married")
```

```{r}
# Create factor version
dt[, married_f := factor(married, labels = c("Not married or intimate relationship", 
                                             "Married or intimate relationship"))]

# Descriptives at day 0
dt[days == 0][order(married_f), .(Women = .N), by = married_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### CTOS

[Study component](https://www.whi.org/researchers/data/WhiDataDict/dem_ctos_inv.pdf)

```{r}
# Create factor version
dt[, ctos_f := factor(ctos, labels = c("CT", "OS"))]

# Descriptives at first observation
dt[obs == 1, .(Women = .N), by = ctos_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```









-------------------------------------------------------------------------------

# Health and health risk factors {#health}

-------------------------------------------------------------------------------

### Smoking status

[Pack years of smoking](https://www.whi.org/researchers/data/WhiDataDict/f34_ctos_inv.pdf)

Pack-years of smoking is a computed variable taking into account years of smoking and number of cigarettes smoked per day on average. Please see the PACKYRS algorithm notes (Pack Years of Smoking Algorithm May 2014.doc) for additional detail on how it is computed.

Over all observations

```{r}
dt[, .(Missing = sum(is.na(packyrs)), Mean = mean(packyrs, na.rm = T), SD = sd(packyrs, na.rm = T))]
```

Pack years of smoking at first observation

```{r}
dt[obs == 1, .(Missing = sum(is.na(packyrs)), Mean = mean(packyrs, na.rm = T), SD = sd(packyrs, na.rm = T))]
```

Pack years of smoking at first observation with a non-missing value

```{r}
dt[!is.na(packyrs), first_packyrs := seq_len(.N) == 1, by = id] # Tag first non-missing value

dt[first_packyrs == 1, 
   .(Missing = sum(is.na(packyrs)), Mean = mean(packyrs, na.rm = T), SD = sd(packyrs, na.rm = T))]
```

### BMI

[Body-mass Index (BMI), kg/m2. If a participant's BMI is greater than 70, BMI is missing.](https://www.whi.org/researchers/data/WhiDataDict/f80_ctos_inv.pdf)

```{r}
dt[, .(Missing = sum(is.na(bmi)), Mean = mean(bmi, na.rm = TRUE), SD = sd(bmi, na.rm = TRUE))]
```

At first observation with a non-missing value

```{r}
dt[!is.na(bmi), first_bmi := seq_len(.N) == 1, by = id] # Tag first non-missing value

dt[first_bmi == 1, 
   .(Missing = sum(is.na(bmi)), Mean = mean(bmi, na.rm = T), SD = sd(bmi, na.rm = T))]
```

Create categorical BMI variable

```{r}
dt <- dt %>% mutate(bmi4cat = case_when(
  is.na(bmi) ~ NA_real_, # Missing
  bmi < 25 ~ 1,          # < 25
  bmi < 30 ~ 2,          # 25-29
  bmi < 35 ~ 3,          # 30-34
  TRUE     ~ 4           # 35+
  )
) %>% 
as.data.table()
```

```{r}
dt[, bmi4cat_f := factor(bmi4cat, labels = c("< 25", "25-29", "30-34", "35+"))]
dt[obs ==1][order(bmi4cat_f), .(Women = .N), by = bmi4cat_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Hysterectomy ever

[Did you ever have a hysterectomy? (This is an operation to take out your uterus or womb.)](https://www.whi.org/researchers/data/WhiDataDict/f2_ctos_inv.pdf)

```{r}
# Create factor version
dt[, hyst_f := factor(hyst, labels = c("No", "Yes"))]

# Descriptives at first observation
dt[obs == 1, .(Women = .N), by = hyst_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Urinary incontinence

[Have you ever leaked even a very small amount of urine involuntarily and you couldn't control it?](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf)

```{r}
# Create factor version
dt[, incont_f := factor(incont, labels = c("No", "Yes"))]

# Tag first non-missing value
dt[!is.na(incont), first_incont := seq_len(.N) == 1, by = id]

# Descriptives at first response
dt[first_incont == 1, .(Women = .N), by = incont_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Hypertension

[Did a doctor ever say that you had hypertension or high blood pressure? (Do not include high
blood pressure that you had only when you were pregnant.)](https://www.whi.org/researchers/data/WhiDataDict/f30_ctos_inv.pdf)

```{r}
# Create factor version
dt[, hypt_f := factor(hypt, labels = c("No", "Yes"))]

# Descriptives at first observation
dt[, .(Obs = .N), by = hypt_f][, Cumsum := cumsum(Obs)][, Percent := Obs / max(Cumsum) * 100][]
```

### Cardiovascular disease

[Has a doctor ever told you that you had heart problems, problems with your blood circulation, or
blood clots?](https://www.whi.org/researchers/data/WhiDataDict/f30_ctos_inv.pdf)

```{r}
# Create factor version
dt[, cvd_f := factor(cvd, labels = c("No", "Yes"))]

# Descriptives at first observation
dt[, .(Obs = .N), by = cvd_f][, Cumsum := cumsum(Obs)][, Percent := Obs / max(Cumsum) * 100][]
```

### Arthritis

[Did your doctor ever say that you had arthritis?](https://www.whi.org/researchers/data/WhiDataDict/f30_ctos_inv.pdf)

```{r}
# Create factor version
dt[, arthrit_f := factor(arthrit, labels = c("No", "Yes"))]

# Descriptives at first observation
dt[, .(Obs = .N), by = arthrit_f][, Cumsum := cumsum(Obs)][, Percent := Obs / max(Cumsum) * 100][]
```

### Diabetes ever

[Did a doctor ever say that you had sugar diabetes or high blood sugar when you were not
pregnant?](https://www.whi.org/researchers/data/WhiDataDict/f2_ctos_inv.pdf)

[Computed from Form 2, question 23, 23.4 and 23.6. Indicator for whether the participant has
ever been treated for diabetes with pills or shots.](https://www.whi.org/researchers/data/WhiDataDict/f2_ctos_inv.pdf)

```{r}
# Create factor version
dt[, diab_f := factor(diab, labels = c("No", "Yes"))]

# Descriptives at first observation
dt[, .(Obs = .N), by = diab_f][, Cumsum := cumsum(Obs)][, Percent := Obs / max(Cumsum) * 100][]
```

Update 2017-09-26: Combine diab and diabtrt

```{r}
dt[, diab_combined := case_when(
  diab == 1 | diabtrt == 1 ~ 1,       # Yes
  diab == 0 & diabtrt == 0 ~ 0,       # No
  TRUE                     ~ NA_real_ # Missing
)]
dt[, diab_combined_f := factor(diab_combined, levels = c(0, 1), labels = c("No", "Yes"))]
dt[, .(Obs = .N), by = diab_combined_f][, Cumsum := cumsum(Obs)][, Percent := Obs / max(Cumsum) * 100][]
```

Checked back to SAS f_2. There aren't actually any instances where a woman said no to diab and yes to diabtrt.

### Cancer

[Did a doctor ever say that you had cancer, a malignant growth, or tumor? (This does not include
"fibroids" of the uterus.)](https://www.whi.org/researchers/data/WhiDataDict/f30_ctos_inv.pdf)

```{r}
# Create factor version
dt[, canc_f := factor(canc_f30, labels = c("No", "Yes"))]

# Descriptives at first observation
dt[, .(Obs = .N), by = canc_f][, Cumsum := cumsum(Obs)][, Percent := Obs / max(Cumsum) * 100][]
```

### Hip fracture after age 55

[Hip fracture age 55 or older](https://www.whi.org/researchers/data/WhiDataDict/f30_ctos_inv.pdf)

```{r}
# Create factor version
dt[, hip55_f := factor(hip55, labels = c("No", "Yes"))]

# Descriptives at first observation
dt[, .(Obs = .N), by = hip55_f][, Cumsum := cumsum(Obs)][, Percent := Obs / max(Cumsum) * 100][]
```

### General health

[In general, would you say your health is (Mark one oval.)](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf)

```{r}
# Create factor version
dt[, genhel_f := factor(genhel, labels = c("Excellent", "Very good", "Good", "Fair", "Poor"))]

# Tag first non-missing value
dt[!is.na(genhel), first_genhel := seq_len(.N) == 1, by = id]

# Descriptives at first response
dt[first_genhel == 1][order(genhel_f), .(Women = .N), by = genhel_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

```{r}
# Dichotomize
dt[genhel_f %in% c("Fair", "Poor"), good_health := 0]
dt[genhel_f %in% c("Excellent", "Very good", "Good"), good_health := 1]

# Factor version
dt[, good_health_f := factor(good_health, labels = c("No", "Yes"))]

# Descriptives at first response
dt[first_genhel == 1, .(Women = .N), by = good_health_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Depression

[Shortened CES-D/DIS Screening Instrument. Computed from Form 36/37, questions 103-108, 109, and 110. Source: Center for Epidemiological Studies; depression scale (CES-D, short form). PSHTDEP ranges from 0 to 1 with a higher score indicating a greater likelihood of depression. Cutoff values of .06 and .009 have been used to indicate depression.](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf)

```{r}
dt[, .(Missing = sum(is.na(pshtdep)), Mean = mean(pshtdep, na.rm = TRUE), SD = sd(pshtdep, na.rm = TRUE))]
```

### Taking SSRI

[Medication Therapeutic Class Code](https://www.whi.org/researchers/data/WhiDataDict/f44_ctos_inv.pdf)

tccode = 581600

```{r}
# Create factor version
dt[, ssri_f := factor(ssri, labels = c("No", "Yes"))]

# Descriptives at first observation
dt[obs == 1][order(ssri_f), .(Women = .N), by = ssri_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```









-------------------------------------------------------------------------------

## Menopausal symptoms and treatment variables {#menopausal}

-------------------------------------------------------------------------------

### Night sweats

[Below is a list of symptoms people sometimes have. For each item, mark the one oval that best
describes how bothersome the symptom was during the past 4 weeks for you. Be sure to mark
one oval on each line. Night sweats](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf)

```{r}
# Create factor version
dt[, nightswt_f := factor(nightswt, labels = c("Symptom did not occur", "Symptom was mild", 
                                               "Symptom was moderate", "Symptom was severe"))]

# Tag first non-missing value
dt[!is.na(nightswt), first_nightswt := seq_len(.N) == 1, by = id]

# Descriptives at first response
dt[first_nightswt == 1, .(Women = .N), by = nightswt_f][, Cumsum := cumsum(Women)][, Percent := Women / max(Cumsum) * 100][]
```

```{r}
# Dichotomize
dt[, night_sweats := if_else(nightswt_f == "Symptom did not occur", 0, 1, NA_real_)]

# Create factor version
dt[, night_sweats_f := factor(night_sweats, labels = c("No", "Yes"))]

# Descriptives
dt[first_nightswt == 1, .(Women = .N), by = night_sweats_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Hot flashes

[Below is a list of symptoms people sometimes have. For each item, mark the one oval that best
describes how bothersome the symptom was during the past 4 weeks for you. Be sure to mark
one oval on each line. Hot flashes](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf)

```{r}
# Create factor version
dt[, hotflash_f := factor(hotflash, labels = c("Symptom did not occur", "Symptom was mild", 
                                               "Symptom was moderate", "Symptom was severe"))]

# Tag first non-missing value
dt[!is.na(hotflash), first_hotflash := seq_len(.N) == 1, by = id]

# Descriptives at first response
dt[first_hotflash == 1][order(hotflash_f), .(Women = .N), by = hotflash_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

```{r}
# Dichotomize
dt[, hot_flashes := if_else(hotflash_f == "Symptom did not occur", 0, 1, NA_real_)]

# Create factor version
dt[, hot_flashes_f := factor(hot_flashes, labels = c("No", "Yes"))]

# Descriptives
dt[first_hotflash == 1, .(Women = .N), by = hot_flashes_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Vaginal dryness

[Below is a list of symptoms people sometimes have. For each item, mark the one oval that best
describes how bothersome the symptom was during the past 4 weeks for you. Be sure to mark
one oval on each line. Vaginal or genital dryness](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf)

```{r}
# Create factor version
dt[, vagdry_f := factor(vagdry, labels = c("Symptom did not occur", "Symptom was mild", 
                                               "Symptom was moderate", "Symptom was severe"))]
# Tag first non-missing value
dt[!is.na(vagdry), first_vagdry := seq_len(.N) == 1, by = id]

# Descriptives at first response
dt[first_vagdry == 1][order(vagdry_f), .(Women = .N), by = vagdry_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

```{r}
# Dichotomize
dt[, vag_dry := if_else(vagdry_f == "Symptom did not occur", 0, 1, NA_real_)]

# Create factor version
dt[, vag_dry_f := factor(vag_dry, labels = c("No", "Yes"))]

# Descriptives
dt[first_vagdry == 1, .(Women = .N), by = vag_dry_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

### Hormones now

[Female hormones now. Are you using female hormones now?](https://www.whi.org/researchers/data/WhiDataDict/f2_ctos_inv.pdf)

```{r}
# Create factor version
dt[, hormnw_f := factor(hormnw, labels = c("No", "Yes"))]

# Tag first non-missing value
dt[!is.na(hormnw), first_hormnw := seq_len(.N) == 1, by = id]

# Descriptives at first response
dt[first_hormnw == 1, .(Women = .N), by = hormnw_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

```{r}
# Clean up
rm(ever_vars, time_invariant)
```









-------------------------------------------------------------------------------

# Save progress {#save}

-------------------------------------------------------------------------------

```{r save_analysis_09}
analysis_09 <- as_tibble(dt)
write_feather(analysis_09, path = "../data/analysis_09.feather")
```

```{r session_info, echo=FALSE}
sessionInfo()
```