---
title: "Preprocess 04: Manage Sexual Function Variables"
date: "Created: 2017-07-17 <br> Updated: `r Sys.Date()`"
output: 
  github_document: default
---

```{r load_packages, message=FALSE}
# Load packages
library(tidyverse)
library(feather)
library(data.table)
library(zoo)
library(gmodels)

# Load functions
source("../R scripts/functions.R")
```

-------------------------------------------------------------------------------

# Clean outcome variables: Sexual function

-------------------------------------------------------------------------------

```{r load_analysis_06}
# Load data
analysis_06 <- read_feather("../data/analysis_06.feather")
check_data(analysis_06) # 1,432,448 observations and 83 variables
```

```{r}
dt <- as.data.table(analysis_06)
```

## Sexual activity in past year

[Did you have any sexual activity with a partner in the last year?](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf)

```{r}
# Recode 9 to missing
dt[sexactiv == 9, sexactiv := NA_real_]
# select(dt, id, days, obs, sexactiv)
```

Many women do not have a measure for sexactiv in their first observation. Below I will create a variable that captures their sexactiv status at the first time it is measured.

```{r}
# Tag first non-missing value of sexactiv
dt[!is.na(sexactiv), first_sexactiv := seq_len(.N) == 1, by = id]
# select(dt, id, days, obs, sexactiv, first_sexactiv)
```

Because of the way the question is worded, we can carry responses backwards within the year. However, this changes the interpretation to be somethin like "where you currently sexually active on this day?"

```{r}
# Temp variable for response number
dt[!is.na(sexactiv), response_num := 1:.N, by = id]

# Temp variable for days at response
dt[!is.na(response_num), response_days := days]

# Carry backward response_days within id
dt[, response_days := na.locf(response_days, na.rm = FALSE, fromLast = TRUE), by = id]

# Absolute difference in days between days and response_days
dt[, diff := abs(response_days - days)]

# Carry backward sexactiv_f if diff <= 365
dt[diff <= 365, sexactiv := na.locf(sexactiv, na.rm = FALSE, fromLast = TRUE), by = id]

# select(dt, id, days, obs, sexactiv, response_num, response_days, diff)

# Drop temp vars
dt[, c("response_num", "response_days", "diff") := NULL]

# Create factor version
dt[, sexactiv_f := factor(sexactiv, labels = c("No", "Yes"))]

# Descriptives at first response
dt[first_sexactiv == 1, .(Women = .N), by = sexactiv_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

```{r}
# How many never respond?
dt[, all(is.na(first_sexactiv)), by = id][, .(`Never respond` = sum(V1))][] # 2,859
```


## Sexual satisfaction

[How satisfied are you with your current sexual activities, either with a partner or alone? (Mark
one oval.)](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf)

```{r}
# Recode 9 to missing
dt[satsex == 9, satsex := NA_real_]

# Create factor version
dt[, satsex_f := factor(satsex, labels = c("Very unsatisfied", "A little unsatisfied", "Somewhat satisfied",
                                           "Very satisfied"))]

# Tag first non-missing value of satsex
dt[!is.na(satsex), first_satsex := seq_len(.N) == 1, by = id]

# Descriptives at first response
dt[first_satsex == 1, .(Women = .N), by = satsex_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

```{r}
# How many never respond?
dt[, all(is.na(first_satsex)), by = id][, .(`Never respond` = sum(V1))][] # 18,064
```

```{r}
# Create a dichotomous version
dt[, satisfied := if_else(satsex_f == "Somewhat satisfied" | satsex_f == "Very satisfied", 1, 0, NA_real_)]

# Create factor version
dt[, satisfied_f := factor(satisfied, labels = c("No", "Yes"))]

# Descriptives at first response
dt[first_satsex == 1, .(Women = .N), by = satisfied_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

## Sexual frequency satisfaction

[Are you satisfied with the frequency of your sexual activity, or would you like to have sex more or
less often? (Mark one oval.)](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf)

```{r}
# Recode 9 to missing
dt[satfrqsx == 9, satfrqsx := NA_real_]

# Create factor version
dt[, satfrqsx_f := factor(satfrqsx, labels = c("Less often", "Satisfied with current frequency", "More often"))]

# Tag first non-missing value of satsex
dt[!is.na(satfrqsx), first_satfrqsx := seq_len(.N) == 1, by = id]

# Descriptives at first response
dt[first_satfrqsx == 1, .(Women = .N), by = satfrqsx_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

```{r}
# How many never respond?
dt[, all(is.na(first_satfrqsx)), by = id][, .(`Never respond` = sum(V1))][] # 26,229
```

```{r}
# Create a dichotomous version
dt[, freq_satisfied := if_else(satfrqsx_f == "Satisfied with current frequency", 1, 0, NA_real_)]

# Create factor version
dt[, freq_satisfied_f := factor(freq_satisfied, labels = c("No", "Yes"))]

# Descriptives at first response
dt[first_satfrqsx == 1, .(Women = .N), by = freq_satisfied_f][, Cumsum := cumsum(Women)][
  , Percent := Women / max(Cumsum) * 100][]
```

```{r save_analysis_07}
# Save progress
analysis_07 <- dt
write_feather(analysis_07, path = "../data/analysis_07.feather")
```

```{r session_info, echo=FALSE}
sessionInfo()
```