---
title: "Preprocess 06: Subset Baseline Data"
date: "Created: 2017-07-17 <br> Updated:`r Sys.Date()`"
output: 
  html_notebook:
    css: custom-css.css
---

# Overview

In this file, we reduce the data to one row per woman (baseline) and to only the variables of interest for our final analysis. 

* Baseline is either each woman's first response to an abuse question or day 0 if she never responded to either abuse question.

This file also serves to create **participant flow diagram** used in the manuscript.

# Table of contents

1. [Subset baseline](#subset-baseline)   
2. [Missing sexual activity](#missing-sexual-activity)    
3. [Missing abuse](#missing-abuse)   
4. [Variable management](#variable-management)   
5. [Save progress](#save-progress)   

```{r load_packages, message=FALSE}
Sys.setenv(TZ = "US/Central")
library(tidyverse)
library(feather)
library(bfuncs)
library(mice)
```

* Primary exposure of interest:     
    + Abuse (Yes / NO)     
    + Abuse (4-level)   
    
* Sexual activity in past year is another exposure of special interest. We are only interested in the association between abuse and sexual function among women that were sexually active.
    
* Outcomes of interest:   
    + Sexual satisfaction   
    + Sexual frequency satisfaction
    
* Covariates of interest:   
    + Age    
    + Race / ethnicity    
    + Education   
    + Income   
    + Marital status    
    + Study component 
    + Smoking 
    + BMI
    + Hysterectomy 
    + Urinary incontinence
    + Hypertension
    + Cardivascular disease
    + Arthritis
    + Diabetes
    + Cancer
    + Hip fracture after age 55
    + Self-rated health
    + Depressive symptoms
    + Taking SSRIs 
    + Night sweats    
    + Hot flashes    
    + Vaginal dryness
    + Hormone use    

* The first set of analyses will be cross-sectional, and use only baseline measures

Currently, all 161,808 women are still represented in the data.

```{r load_analysis_09}
analysis_09 <- read_feather("../data/analysis_09.feather")
about_data(analysis_09) # 1,432,448 observations and 136 variables
```









-------------------------------------------------------------------------------

# Subset baseline

-------------------------------------------------------------------------------

## Tag baseline observations

Baseline is either each woman's first response to an abuse question or day 0 if she never responded to either abuse question.

* first_ab_obs was created in preprocess_03_abuse.Rmd. It is a dummy variable that represent the first time that there is a non-missing value for the physical abuse question or the verbal abuse question.

* abuse_ever was created in preprocess_03_abuse.Rmd. abuse_ever == -3 means that the current woman never had a non-missing value for the physical abuse question or the verbal abuse question at any point during follow-up.

```{r}
baseline <- analysis_09 %>% 
  mutate(
    baseline = case_when(
      first_ab_obs == TRUE         ~ 1L, # First observed value for abuse
      abuse_ever == -3 & days == 0 ~ 1L, # Day 0, if never observed value for abuse
      TRUE                         ~ 0L  # Otherwise, not baseline row
    )
  )

about_data(baseline) # 1,432,448 observations and 137 variables
```

Count baseline observations

```{r}
baseline %>% group_by(baseline) %>% freq_table()
```

As expected, there are 161,808 baseline rows.

## Keep baseline observations only

```{r}
baseline_only <- baseline %>% 
  filter(baseline == 1)

about_data(baseline_only) # 161,808 observations and 137 variables
```

## What is the distribution of CT/OS for flow diagram?

```{r}
baseline_only %>% 
  group_by(ctos_f) %>% 
  freq_table()
```

## Subset the variables of interest

2017-06-26: Added hypertension
2018-01-17: Added inc2cat_f

```{r subset_vars_of_interest}
baseline_only <- baseline_only %>% 
  select(
    id, days, abuse_d_f, abuse4cat_f, sexactiv_f, satisfied_f, freq_satisfied_f, satfrqsx_f,
    age, age_group_f, race_eth_f, edu4cat_f, edu2cat_f, inc5cat_f, inc2cat_f, married_f, ctos_f, 
    packyrs, bmi, bmi4cat_f, hyst_f, incont_f, hypt_f, cvd_f, arthrit_f, diab_f, 
    diab_combined_f, canc_f, hip55_f, good_health_f, pshtdep, ssri_f, first_nightswt, 
    hot_flashes_f, night_sweats_f, vag_dry_f, hormnw_f, sex, baseline
  )

about_data(baseline_only) # 161,808 observations and 39 variables
```









-------------------------------------------------------------------------------

# Missing sexual activity

-------------------------------------------------------------------------------

There are two relevant questions here.

* From [form 37](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf) and [form 38](https://www.whi.org/researchers/data/WhiDataDict/f38_ctos_fu_inv.pdf): Did you have any sexual activity with a partner in the last year?

* From [form 37](https://www.whi.org/researchers/data/WhiDataDict/f37_ctos_inv.pdf): Regardless of whether you are currently sexually active, which response best describes who you have had sex with over your adult lifetime?

Remove women who say they haven't had sex in the past year and women who say that they have never had sex. Additionally, I need to remove women who are missing this information. I can't just impute a value because I am sub-setting (dropping women who were not sexually active in the past year) based on this variable, as opposed to modeling. Imputing will potentially give me different values for sexual activity in the past year within woman.

```{r rows.print=18}
baseline_only %>% 
  group_by(sexactiv_f, sex) %>% 
  freq_table()
```

There are 73 women who report never having sex and having sex in the past year. To be conservative, we will drop these women.

```{r}
had_sex <- baseline_only %>% 
  mutate(
    had_sex = case_when(
      sexactiv_f == "No" | sex == 1 ~ FALSE,
      sexactiv_f == "Yes"           ~ TRUE,
      TRUE                          ~ NA
    )
  )

had_sex %>% 
  group_by(had_sex) %>% 
  freq_table()
```

```{r}
71371 + 7108
```

```{r}
had_sex <- had_sex %>% filter(had_sex)
about_data(had_sex) # 83,329 observations and 40 variables
```

```{r}
had_sex %>% 
  group_by(ctos_f) %>% 
  freq_table()
```

71,371 reported not having sex in the past year and the status of past year sexual activity could not be determined for 7,108. Those 78,479 women were excluded from the data -- **leaving 83,329 women remaining for analysis.** 









-------------------------------------------------------------------------------

# Missing abuse

-------------------------------------------------------------------------------

Originally, I dropped women with missing abuse and missing sexual function. Now, we will use MICE to impute values for abuse and sexual function instead.

For abuse, we will impute the 4-level abuse variable, and then recode to the 2-level abuse variable again within each imputed dataset.

## How many women are missing 4-level abuse at baseline?

```{r}
had_sex %>% 
  group_by(abuse4cat_f) %>% 
  freq_table()
```

```{r}
had_sex %>% 
  group_by(abuse_d_f) %>% 
  freq_table()
```

Of the 83,329 women that reported sexual activity in the past year, 2 never answered either abuse question and 182 women's abuse status could not be determined from their answers. The abuse status (4-level) can be determined for over 99% of the 83,329 women who reported sexual activity in the past year.

Note that if we use any abuse vs. no abuse instead of the 4-category abuse variable, then there are fewer missing (160). However, because we want the ability to investigate the importance of the 4 category levels in subsequent analyses, and we are imputing values anyway, we will use the count of missing values from abuse4cat_f in the **participant flow diagram**.

### 2018-01-17 update

After conducting MI and fitting the regression models to the data, we decided to use the 2-level abuse variable as our exposure. Therefore, we will now use the count of missing values from abuse_d_f (**160**) in the **participant flow diagram**.

## Overview of data for Table 1 - for participant flow chart

```{r}
had_sex %>% 
  filter(abuse_d_f == "No" | abuse_d_f == "Yes") %>% 
  group_by(ctos_f) %>% 
  freq_table()
```









-------------------------------------------------------------------------------

# Variable management

-------------------------------------------------------------------------------

## Basline only data for Table 1

```{r}
baseline_only <- had_sex
about_data(baseline_only) # 83,329 observations and 40 variables
```

## Baseline only data for multiple imputation and modeling

```{r}
baseline_only_mi <- had_sex %>%
  select(
    # Administrative variables
    # Calling CTOS administrative. There is no difference in satisfaction by CTOS
    id, days, baseline, ctos_f,
    
    # Response variables
    satisfied_f, freq_satisfied_f,
    
    # Abuse variables
    # Since we are only using 4-level abuse in the MI models, go ahead and drop 
    # the 2-level abuse variable.
    abuse4cat_f,
    
    # Sociodemographic covariates
    age, race_eth_f, edu2cat_f, inc5cat_f, married_f,
    
    # Health and health risk factors
    packyrs, bmi, hyst_f, incont_f, hypt_f, cvd_f, arthrit_f, diab_f, canc_f, hip55_f, good_health_f,
    pshtdep, ssri_f, 
    
    # Menopausal symptoms and treatment variables
    night_sweats_f, hot_flashes_f, vag_dry_f, hormnw_f
  )

about_data(baseline_only_mi) # 83,329 observations and 29 variables
```

## Recode abuse4cat to NA

I also, need to recode indicator values for different kinds of missing for abuse4cat_f to NA

```{r}
baseline_only_mi$abuse4cat_f <- recode_factor(
  baseline_only_mi$abuse4cat_f, 
  "Can't determine V no P" = NA_character_,
  "Can't determine P no V" = NA_character_,
  "Abuse not measured" = NA_character_
)
```

## Create imputation flags

```{r}
ini <- mice(baseline_only_mi, m = 0, maxit = 0)

# For imputation flags, only need vars with any missing data.
vars_w_missing <- ini$nmis[ini$nmis > 0]
vars_w_missing <- names(vars_w_missing)

vars_w_missing
```

For each member of vars_w_missing create a impute_ flag variable

```{r}
for (var in vars_w_missing) {
  var_name <- paste0("impute_", var)
  baseline_only_mi[[var_name]] <- if_else(is.na(baseline_only_mi[[var]]), 1, 0)
}

# names(baseline_only)
# baseline_only %>% select(id, diab_f, impute_diab_f)
about_data(baseline_only_mi) # 83,329 observations and 53 variables
```

Clean up

```{r}
rm(analysis_09, baseline, had_sex, ini, var, var_name, vars_w_missing)
```









-------------------------------------------------------------------------------

# Save progress

-------------------------------------------------------------------------------

## Save feather files

```{r}
write_feather(baseline_only, "../data/baseline_only.feather")
```

```{r}
write_feather(baseline_only_mi, "../data/baseline_only_mi.feather")
```

## Export as csv

```{r}
readr::write_csv(baseline_only_mi, "../data/baseline_only_mi.csv")
```

```{r session_info, echo=FALSE}
sessionInfo()
```