---
title: "Preprocess 07: Multiple Imputation"
date: "Created: 2017-05-29 <br> Updated: `r Sys.Date()`"
output: html_notebook
discussion: Do this part last. Describe what actually happens in this file.
---

<style>
  hr {
    background-color: #66b3ff;
    height: 1px;
  }
</style>

```{r load_packages, message=FALSE}
Sys.setenv(TZ = "US/Central")

# Load packages
library(feather)
library(tidyverse)
library(data.table)
library(mice)
library(lubridate)
library(bfuncs)
```

-------------------------------------------------------------------------------

# Preprocessing for Multiple Imputation

-------------------------------------------------------------------------------

```{r load_baseline_only}
# Load data
baseline_only <- read_feather("../data/baseline_only.feather")
about_data(baseline_only) # 83,329 observations and 53 variables
```

```{r}
count_ids(baseline_only$id) # 83,329 unique values
```

Currently, the data includes information from 83,145 unique women. All women reported sexual activity in the past year. Additionally, each woman's abuse status can be determined.

## View the number and percent missing for each variable

```{r}
ini <- mice(baseline_only, maxit = 0)
```

```{r}
ini$nmis[order(ini$nmis)]
```

```{r}
(ini$nmis[order(ini$nmis)] / nrow(baseline_only) * 100) %>% round(1)
```

* Lots of missing (8.5% to 99.1%) for: canc_f, packyrs, cvd_f, hypt_f, hip55_f, arthrit_f, hyst_f, bmi, bmi4cat_f, ssri_f, diab_f, diab_combined_f, and hormnw_f

## Add the age/abuse interaction

Buuren, S., & Groothuis-Oudshoorn, K. (2011). mice: Multivariate imputation by chained equations in R. Journal of Statistical Software. Retrieved from http://doc.utwente.nl/78938/

View the dummy terms for categorical variables

```{r}
ini$pad$data %>% select(contains("abuse"))
```

The factor abuse4cat is internally represented by dummy variables abuse4cat_f.1, abuse4cat_f.2, and abuse4cat_f.3. The interaction between age and abuse can be added as:

```{r}
baseline_only_w_interaction <- baseline_only %>% 
  mutate(
    age_abuse_phys_only = NA_integer_,
    age_abuse_verb_only = NA_integer_,
    age_abuse_both      = NA_integer_
  )
```

Create a mean centered version of age before creating the interaction term

```{r}
baseline_only_w_interaction <- baseline_only_w_interaction %>% 
  mutate(age_c = age - mean(age))
```

```{r}
ini <- mice(baseline_only_w_interaction, m = 0, maxit = 0)
```

Adjust MI methods

```{r}
meth <- ini$method
meth["age_abuse_phys_only"] <- "~I(abuse4cat_f.1 * age)"
meth["age_abuse_verb_only"] <- "~I(abuse4cat_f.2 * age)"
meth["age_abuse_both"]      <- "~I(abuse4cat_f.3 * age)"
```

Adjust prediction matrix again

* Make sure administrative variables are not being used for imputation (e.g. id, ctos)

* Don't predict main effect terms from interaction terms

```{r}
pred <- ini$predictorMatrix 

pred[, c("id", "days", "ctos_f")] <- 0

pred[c("age", "abuse4cat_f"), 
     c("age_abuse_phys_only", "age_abuse_verb_only", "age_abuse_both")] <- 0
```






-------------------------------------------------------------------------------

# Adjust multiple imputation settings

-------------------------------------------------------------------------------

## View logged events

```{r logged_events}
ini <- mice(baseline_only, m = 0, maxit = 0) # Dry run to view information about the imputation process
ini$loggedEvents
```

Need to drop the baseline variable

```{r}
baseline_only <- baseline_only %>% select(-baseline)
about_data(baseline_only) # 83,329 observations and 52 variables
```

Check again

```{r}
ini <- mice(baseline_only, m = 0, maxit = 0)
ini$loggedEvents
```

Order data by missingness to speed up imputations

```{r}
order <- ini$nmis
order <- sort(order)
order <- names(order)
baseline_only <- baseline_only %>% select(order)
```




## Adjust prediction matrix   

* Copy the predictor matrix to a new object   
* Easiest to view the pred matrix in data viewer    

```{r}
ini <- mice(baseline_only, m = 0, maxit = 0)
pred <- ini$predictorMatrix 
```

The _rows_ of the predictor matrix correspond to the variables being predicted (or not). The _columns_ of the predictor matrix correspond to the variables being used to predict.

* All completely observed vars (e.g., age) are not predicted by default (i.e., row of 0's)

* Make sure administrative variables are not being used for imputation (e.g. id, ctos)   

* Make sure var_days variables aren't imputing anything other than their specific var. For example, bmi_days should only be used to impute bmi, not marital status. __Update: we dropped var_days__

* I don't think we want to impute the "_2" or "_days" variables. We are just using those to help predict baseline values for their respective vars. If the var (e.g. canc_f) is missing for every row within an observation, we still want to impute that value based on other variables (age, general health, etc.), but there isn't much to be gained by also imputing the "next closet value" for that variable.

* Adjust any other variables that appear to be noise.   

* Set cells of the matrix to 0 or 1 as needed, using index [r, c] notation

### Don't use administrative variables to impute anything

```{r}
pred[, c("id", "days", "ctos_f")] <- 0
```




## Adjust MI methods

* To prevent vars from being imputed, we need to set the method to missing in addition to adjusting the predictor matrix.   

```{r}
meth <- ini$method
meth
```









-------------------------------------------------------------------------------

# Generate the imputations

-------------------------------------------------------------------------------

## No interaction terms

```{r}
start <- now()

imp <- mice(data = baseline_only,
            m = 10,                     # Default is 5
            method = meth,              # Custom method vector created above
            predictorMatrix = pred,     # Custom predictor matrix created above
            seed = 20170511)            # For reproducability

stop <- now()
time <- difftime(stop, start)
```

```{r}
time # MI took 2.598068 hours
```

## Save Multiply imputed data set (MIDS) object from mice

```{r}
imp_83329 <- imp
write_rds(imp_83329, path = "../data/imp_83329.rds")
```

## Added interaction (2018-01-13)

```{r}
start <- now()

imp_interaction <- mice(data = baseline_only_w_interaction,
            m = 10,                     # Default is 5
            method = meth,              # Custom method vector created above
            predictorMatrix = pred,     # Custom predictor matrix created above
            seed = 20170511)            # For reproducability

stop <- now()
time <- difftime(stop, start)
```

```{r}
time # MI took 2.673914 hours
```

## Save Multiply imputed data set (MIDS) object from mice

```{r}
write_rds(imp_interaction, path = "../data/imp_interaction.rds")
```

# Stack the imputed datasets

```{r}
analysis_11_interaction <- complete(imp_interaction, action = "long", include = TRUE)
```

How many obs and vars in each imputation set?

```{r}
analysis_11_interaction %>% 
  filter(.imp == 1) %>% 
  about_data() # 83,329 observations and 57 variables
```

Check for missing

Overall:

```{r}
analysis_11_interaction %>% 
  map_df(~ sum(is.na(.))) %>% 
  gather(key = variable, value = missing)
```

Because I used "include = TRUE" above, the stacked data includes a copy of the original data with missing. I needed to do this for the as.mids function below to work.

Imputed data sets only:

```{r}
analysis_11_interaction %>% 
  filter(!.imp == 0) %>% 
  map_df(~ sum(is.na(.))) %>% 
  gather(key = variable, value = missing)
```

No missing values remain in the imputed data sets.

## Put dichotomous abuse variable back in the data

+ Make a numeric version of satisfied and freq satisfied for modeling.

+ Create dissatisfied variables that will be used for modeling.

Need to change the order of factor levels for abuse4cat_f. Currently: "Physical and verbal abuse" "Physical abuse only" "Verbal abuse only" "Did not experience abuse"

```{r}
analysis_11_interaction <- analysis_11_interaction %>% 
  mutate(
    abuse = case_when(
      abuse4cat_f == "Did not experience abuse" ~ 0,
      abuse4cat_f == "Verbal abuse only" ~ 1,
      abuse4cat_f == "Physical abuse only" ~ 1,
      abuse4cat_f == "Physical and verbal abuse" ~ 1
    ),
    satisfied = if_else(satisfied_f == "Yes", 1L, 0L, NA_integer_),
    freq_satisfied = if_else(freq_satisfied_f == "Yes", 1L, 0L, NA_integer_),
    dissatisfied = if_else(satisfied ==1, 0L, 1L, NA_integer_),
    freq_dissatisfied = if_else(freq_satisfied == 1, 0L, 1L, NA_integer_),
    
    abuse4cat_f = forcats::fct_relevel(abuse4cat_f, "Did not experience abuse", "Verbal abuse only",
                                       "Physical abuse only", "Physical and verbal abuse")
  )
```




# Left off here... Need to check to see if the impuations of the interaction terms make sense...





Convert analysis_11 back to a mids object that can be used for modeling: https://cran.r-project.org/web/packages/mice/mice.pdf

```{r}
mids_interaction <- as.mids(analysis_11_interaction)
```

## Save new mids object

```{r}
write_rds(mids_83329, "../data/mids_83329.rds")
```

## Save stacked data in feather format

```{r}
write_feather(analysis_11, "../data/analysis_11.feather")
```

## Export to excel for SAS

```{r}
readr::write_csv(analysis_11, "../data/analysis_11.csv") # Too large to open in Excel.
```

Ideally:

* Need to run descriptive in SAS. Proc Freq, etc.

* See if I can replicate in R. 

* If not, post on SO.

* Run models in SAS MI analyze. RR models vs. OR models.

* See if I can replicate in R.

In reality, we kind of just need to get this done. I think they will want OR models even if OR models are technically correct. For now, just give them table 2. We can always come back and revisit this if we need to.

-------------------------------------------------------------------------------

```{r session_info, echo=FALSE}
# sessionInfo()
```
