---
title: "Preprocess 07: Multiple Imputation"
date: "Created: 2017-05-29 <br> Updated: `r Sys.Date()`"
output: html_notebook
---

<style>
  hr {
    background-color: #66b3ff;
    height: 1px;
  }
</style>

# Overview

In this file, we start with baseline only data for all women who were sexually active in the previous year. We impute values for women who are missing an observed value for abuse (4 category) and/or an observed value for any of the covariates of interest. We perform the imputations using the MICE algorithm (Van Buuren, 2011).

Van Buuren, S., & Groothuis-Oudshoorn, K. (2011). mice: Multivariate imputation by chained equations in R. Journal of Statistical Software. Retrieved from http://doc.utwente.nl/78938/

# Table of contents

1. [Preprocessing for Multiple Imputation](#pre)   
2. [Adjust multiple imputation settings](#settings)    
3. [Generate the imputations](#mi)   
4. [Post processing of imputed data](#post)

```{r load_packages, message=FALSE}
Sys.setenv(TZ = "US/Central")
library(feather)
library(tidyverse)
library(mice)
library(lubridate)
library(bfuncs)
```

-------------------------------------------------------------------------------

# Preprocessing for Multiple Imputation {#pre}

-------------------------------------------------------------------------------

```{r load_baseline_only}
# Load data
baseline_only <- read_feather("../data/baseline_only.feather")
about_data(baseline_only) # 83,329 observations and 53 variables
```

```{r}
count_ids(baseline_only$id) # 83,329 unique values
```

Currently, the data includes information from 83,145 unique women. All women reported sexual activity in the past year. Additionally, each woman's abuse status can be determined.

## View the number and percent missing for each variable

```{r}
ini <- mice(baseline_only, maxit = 0)
```

```{r}
ini$nmis[order(ini$nmis)]
```

```{r}
(ini$nmis[order(ini$nmis)] / nrow(baseline_only) * 100) %>% round(1)
```

* Lots of missing (8.5% to 99.1%) for: canc_f, packyrs, cvd_f, hypt_f, hip55_f, arthrit_f, hyst_f, bmi, bmi4cat_f, ssri_f, diab_f, diab_combined_f, and hormnw_f

## Add the age/abuse interaction {#interaction}

Buuren, S., & Groothuis-Oudshoorn, K. (2011). mice: Multivariate imputation by chained equations in R. Journal of Statistical Software. Retrieved from http://doc.utwente.nl/78938/

View the dummy terms for categorical variables

```{r}
ini$pad$data %>% select(contains("abuse"))
```

The factor abuse4cat is internally represented by dummy variables abuse4cat_f.1, abuse4cat_f.2, and abuse4cat_f.3. The interaction between age and abuse can be added as:

```{r}
baseline_only_w_interaction <- baseline_only %>% 
  mutate(
    age_abuse_phys_only = NA_integer_,
    age_abuse_verb_only = NA_integer_,
    age_abuse_both      = NA_integer_
  )
```









-------------------------------------------------------------------------------

# Adjust multiple imputation settings {#settings}

-------------------------------------------------------------------------------

## View logged events

Dry run to view information about the imputation process

```{r}
ini <- mice(baseline_only_w_interaction, m = 0, maxit = 0)
ini$loggedEvents
```

We need to drop the baseline variable. Don't worry about the interaction variables. They are currently all "NA." They will be filled in during the imputation process.

```{r}
baseline_only_w_interaction <- baseline_only_w_interaction %>% select(-baseline)
about_data(baseline_only_w_interaction) # 83,329 observations and 55 variables
```

Check again

```{r}
ini <- mice(baseline_only_w_interaction, m = 0, maxit = 0)
ini$loggedEvents
```

Order data by missingness to speed up imputations

```{r}
order <- ini$nmis
order <- sort(order)
order <- names(order)
baseline_only_w_interaction <- baseline_only_w_interaction %>% select(order)
rm(order)
```

## Adjust prediction matrix   

* Copy the predictor matrix to a new object   
* Easiest to view the pred matrix in data viewer    

```{r}
ini <- mice(baseline_only_w_interaction, m = 0, maxit = 0)
pred <- ini$predictorMatrix 
```

The _rows_ of the predictor matrix correspond to the variables being predicted (or not). The _columns_ of the predictor matrix correspond to the variables being used to predict.

* All completely observed vars (e.g., age) are not predicted by default (i.e., row of 0's)

* Make sure administrative variables are not being used for imputation (e.g. id, ctos)   

* Adjust any other variables that appear to be noise.   

* Set cells of the matrix to 0 or 1 as needed, using index [r, c] notation

### Don't use administrative variables to impute anything

```{r}
pred[, c("id", "days", "ctos_f")] <- 0
```

### Don't predict main effect terms from interaction terms

```{r}
pred[c("age", "abuse4cat_f"), 
     c("age_abuse_phys_only", "age_abuse_verb_only", "age_abuse_both")] <- 0
```

## Adjust MI methods

* To prevent vars from being imputed, we need to set the method to missing in addition to adjusting the predictor matrix.   

```{r}
meth <- ini$method
meth
```

### Adjust MI methods - move this down

```{r}
meth["age_abuse_phys_only"] <- "~I(abuse4cat_f.1 * age)"
meth["age_abuse_verb_only"] <- "~I(abuse4cat_f.2 * age)"
meth["age_abuse_both"]      <- "~I(abuse4cat_f.3 * age)"

meth
```









-------------------------------------------------------------------------------

# Generate the imputations {#mi}

-------------------------------------------------------------------------------

2018-01-13: Added interaction

```{r}
start <- now()

imp_interaction <- mice(data = baseline_only_w_interaction,
            m = 10,                     # Default is 5
            method = meth,              # Custom method vector created above
            predictorMatrix = pred,     # Custom predictor matrix created above
            seed = 20170511)            # For reproducability

stop <- now()
time <- difftime(stop, start)
```

```{r}
time # MI took 2.673914 hours
```

## Save Multiply imputed data set (MIDS) object from mice

```{r}
write_rds(imp_interaction, path = "../data/imp_interaction.rds")
```









-------------------------------------------------------------------------------

# Post processing of imputed data {#post}

-------------------------------------------------------------------------------

```{r}
imp_interaction <- read_rds(path = "../data/imp_interaction.rds")
```

## Stack the imputed datasets

```{r}
analysis_11_interaction <- complete(imp_interaction, action = "long", include = TRUE)
```

How many obs and vars in each imputation set?

```{r}
analysis_11_interaction %>% 
  filter(.imp == 1) %>% 
  about_data() # 83,329 observations and 57 variables
```

## Check for missing

Overall:

```{r}
analysis_11_interaction %>% 
  map_df(~ sum(is.na(.))) %>% 
  gather(key = variable, value = missing)
```

Because I used "include = TRUE" above, the stacked data includes a copy of the original data with missing. I needed to do this for the as.mids function below to work.

Imputed data sets only:

```{r}
analysis_11_interaction %>% 
  filter(!.imp == 0) %>% 
  map_df(~ sum(is.na(.))) %>% 
  gather(key = variable, value = missing)
```

No missing values remain in the imputed data sets.

Did the interaction terms get imputed as expected?

```{r}
analysis_11_interaction %>% 
  filter(.imp == 1) %>% 
  select(age, abuse4cat_f, age_abuse_phys_only:age_abuse_both)
```

The interaction terms did not run as expected. Need to rerun. Actually, the imputations are fine - I just need to change the variable name so that it makes sense.

```{r}
analysis_11_interaction <- analysis_11_interaction %>% 
  mutate(age_abuse_none = age_abuse_both) %>% 
  select(-age_abuse_both)
```

```{r}
analysis_11_interaction %>% 
  filter(.imp == 1) %>% 
  select(age, abuse4cat_f, age_abuse_phys_only:age_abuse_none)
```

## Put dichotomous abuse variable back in the data

* Make a numeric version of satisfied and freq satisfied for modeling.   
* Create dissatisfied variables that will be used for modeling.   

Need to change the order of factor levels for abuse4cat_f. Currently: "Physical and verbal abuse" "Physical abuse only" "Verbal abuse only" "Did not experience abuse"

```{r}
analysis_11_interaction <- analysis_11_interaction %>% 
  mutate(
    abuse = case_when(
      abuse4cat_f == "Did not experience abuse" ~ 0,
      abuse4cat_f == "Verbal abuse only" ~ 1,
      abuse4cat_f == "Physical abuse only" ~ 1,
      abuse4cat_f == "Physical and verbal abuse" ~ 1
    ),
    satisfied = if_else(satisfied_f == "Yes", 1L, 0L, NA_integer_),
    freq_satisfied = if_else(freq_satisfied_f == "Yes", 1L, 0L, NA_integer_),
    dissatisfied = if_else(satisfied ==1, 0L, 1L, NA_integer_),
    freq_dissatisfied = if_else(freq_satisfied == 1, 0L, 1L, NA_integer_),
    
    abuse4cat_f = forcats::fct_relevel(abuse4cat_f, "Did not experience abuse", "Verbal abuse only",
                                       "Physical abuse only", "Physical and verbal abuse")
  )
```

## Convert analysis_11 back to a mids object that can be used for modeling: 

https://cran.r-project.org/web/packages/mice/mice.pdf

```{r}
mids_interaction <- as.mids(analysis_11_interaction)
```

## Save new mids object

```{r}
write_rds(mids_interaction, "../data/mids_interaction.rds")
```

## Save stacked data in feather format

```{r}
write_feather(analysis_11, "../data/analysis_11.feather")
```

## Export to excel for SAS

```{r}
readr::write_csv(analysis_11_interaction, "../data/analysis_11_interaction.csv")
```

-------------------------------------------------------------------------------

```{r session_info, echo=FALSE}
sessionInfo()
```
